### 2024-10-23 paper/paper.qmd
update paper

---
title: "Election 2024"
author: "Alexander Guarasci & Jacob Gilbert"
date: today
date-format: long
bibliography: references.bib
execute:
  echo: false
format: pdf
fig-pos: "H"
abstract: |

 This paper is about the election
---

# 1 Introduction

# References

### 2024-10-23 scripts/04-test_analysis_data.py
create tests for the real data that is being used in our analysis

@@ -1,0 +1,80 @@
+#### Preamble ####
+# Purpose: Tests... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 26 September 2024 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+import pandas as pd
+
+# Load the dataset
+df = pd.read_csv('data/02-analysis_data/merged_swing_state_data.csv')
+
+def test_trump_harris_pct_sum():
+    """
+    Test to ensure that the sum of Trump and Harris polling percentages does not exceed 100%.
+    """
+    invalid_rows = df[df['trump_pct'] + df['harris_pct'] > 100]
+    assert invalid_rows.empty, f"Found rows where Trump + Harris percentage exceeds 100%: {invalid_rows}"
+
+def test_valid_states():
+    """
+    Test to ensure that only certain states are present in the dataset.
+    """
+    valid_states = ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 'Pennsylvania', 'Wisconsin']
+    invalid_states = df[~df['state'].isin(valid_states)]
+    assert invalid_states.empty, f"Found rows with invalid states: {invalid_states['state'].unique()}"
+
+def test_at_least_one_data_point_per_state():
+    """
+    Test to ensure that there is at least one data point for each of the key states.
+    """
+    required_states = ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 'Pennsylvania', 'Wisconsin']
+    missing_states = [state for state in required_states if state not in df['state'].unique()]
+    assert not missing_states, f"Missing data points for states: {missing_states}"
+
+def test_no_negative_percentages():
+    """
+    Test to ensure that polling percentages (Trump and Harris) are non-negative.
+    """
+    invalid_rows = df[(df['trump_pct'] < 0) | (df['harris_pct'] < 0)]
+    assert invalid_rows.empty, f"Found rows with negative percentages: {invalid_rows}"
+
+def test_no_future_dates():
+    """
+    Test to ensure that start_date and end_date are not in the future.
+    """
+    df['start_date'] = pd.to_datetime(df['start_date'], errors='coerce')
[…]

### 2024-10-23 scripts/xandermodel1.py
add the initial model as well as its outputs

@@ -1,0 +1,64 @@
+# Import necessary libraries
+import pandas as pd
+from sklearn.linear_model import LogisticRegression
+from sklearn.preprocessing import StandardScaler
+import joblib
+
+# Load the cleaned polling data
+data_path = "data/02-analysis_data/merged_swing_state_data.csv"
+clean_poll_data = pd.read_csv(data_path)
+
+# Check if necessary columns exist
+required_columns = ['trump_pct', 'harris_pct', 'end_date', 'sample_size', 'state']
+for column in required_columns:
+    if column not in clean_poll_data.columns:
+        raise ValueError(f"The required column '{column}' is not present in the dataset.")
+
+# Convert end_date column to datetime format
+clean_poll_data['end_date'] = pd.to_datetime(clean_poll_data['end_date'])
+
+# Define a weighting function
+def calculate_weight(row):
+    # Weight based on recency (more recent = higher weight)
+    recency_weight = 1 / (pd.Timestamp.now() - row['end_date']).days  # More recent polls have higher weights
+    sample_size_weight = row['sample_size'] / clean_poll_data['sample_size'].max()  # Scale by max sample size
+    return recency_weight * sample_size_weight
+
+# Apply the weighting function to create a new 'weight' column
+clean_poll_data['weight'] = clean_poll_data.apply(calculate_weight, axis=1)
+
+# Aggregate polling percentages for Trump and Harris by state, weighted by sample size and recency
+state_average = clean_poll_data.groupby('state').agg(
+    trump_pct=('trump_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
+    harris_pct=('harris_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
+    total_weight=('weight', 'sum')  # Total weight for reference
+).reset_index()
+
+# Create a binary outcome for prediction: 1 if Trump is predicted to win, 0 if Harris
+state_average['winner'] = (state_average['trump_pct'] > state_average['harris_pct']).astype(int)
+
+# Features and target
+X = state_average[['trump_pct', 'harris_pct']]
+y = state_average['winner']
+
+# Initialize the Logistic Regression model with L2 regularization
+logistic_model = LogisticRegression(penalty='l2', solver='lbfgs', max_iter=2000)
+
[…]

### 2024-10-23 scripts/xandermodel.py
Commit a new model

@@ -1,0 +1,44 @@
+import pandas as pd
+from sklearn.linear_model import LogisticRegression
+
+# Load the cleaned polling data
+clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")
+
+# Create a weight for each poll based on recency and sample size
+# For simplicity, let's assume that the weight is based on the sample size directly
+# and a recency factor that gives more weight to recent polls
+# You can customize the recency weight according to your requirements
+current_date = pd.to_datetime('today')
+clean_poll_data['start_date'] = pd.to_datetime(clean_poll_data['start_date'])
+clean_poll_data['recency_weight'] = (current_date - clean_poll_data['start_date']).dt.days
+
+# Normalize the recency weight (larger weight for more recent polls)
+clean_poll_data['recency_weight'] = clean_poll_data['recency_weight'].max() - clean_poll_data['recency_weight']
+clean_poll_data['weight'] = clean_poll_data['sample_size'] * clean_poll_data['recency_weight']
+
+# Aggregate polling percentages for Trump and Harris by state, weighted by sample size and recency
+state_average = clean_poll_data.groupby('state').agg(
+    trump_pct=('trump_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
+    harris_pct=('harris_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum())
+).reset_index()
+
+# Calculate the probabilities of Trump winning based on the weighted percentages
+state_average['trump_win_prob'] = state_average['trump_pct'] / (state_average['trump_pct'] + state_average['harris_pct'])
+
+# Prepare the logistic regression model
+X = state_average[['trump_pct', 'harris_pct']]
+y = (state_average['trump_pct'] > state_average['harris_pct']).astype(int)  # Binary outcome for the model
+
+# Fit the logistic regression model
+logistic_model = LogisticRegression()
+logistic_model.fit(X, y)
+
+# Predict probabilities of Trump winning
+state_average['predicted_win_prob'] = logistic_model.predict_proba(X)[:, 1]
+
+# Output the results
+final_results = state_average[['state', 'trump_pct', 'harris_pct', 'predicted_win_prob']].copy()
[…]

### 2024-10-23 paper/paper.qmd
add initial paper.qmd

@@
 
+This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris.
+By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump or Harris victory in each state.
+This approach allows us to analyze voter support patterns across the states and predict election outcomes more accurately.
+
+The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, derived from aggregated state-level polling averages.
+The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, with predictor variables comprising the weighted average polling percentages for both candidates.
+We assign greater importance to more recent polls and those with larger sample sizes, enhancing the reliability of our estimates.
+
+Our model employs a logistic regression framework to predict election outcomes based on these weighted averages.
+The results reveal the geographic distribution of support for each candidate.
+We have focused our analysis on the swing states where polling percentages are closely contested and could significantly influence the final election result.
+
+Accurate election predictions provide valuable insights into voter dynamics, helping political analysts, campaigns, and the public anticipate electoral outcomes.
+By focusing on high-quality polling data and incorporating weights for recency and sample size, our model enhances prediction reliability and identifies key regions where voter sentiment may shift, ultimately impacting the election..
+
+The structure of the paper is as follows
+
+# Data {#sec-data}
+
+The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
+The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages...
+
+# 2.1 Measurement 
+The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
+This dataset reflects potential outcomes for the 2024 U.S.
+Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
+Each entry in the dataset corresponds to a specific poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+
+It’s important to note that each pollster employs unique sampling techniques and methodologies, which influence their numeric grades and transparency scores.
+To ensure the integrity of our analysis, we selected only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
+This approach is used so that the data reflects a robust and credible representation of voter preferences.
+
+The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
+By aggregating these individual responses, we are able to derive state-level averages for both candidates.
+This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+
+ # 2.1 The Dataset
+
+ # 2.2 Variables of Interest
+
+ # 3 Model {#sec-model}
+
+This paper develops a logistic regression model to predict the likelihood of Donald Trump winning various states in the upcoming 2024 U.S.
+Presidential Election based on aggregated polling data.
+The model leverages polling percentages, recency of the polls, and sample sizes to provide a robust probability estimate of Trump’s chances against Kamala Harris.
+
+Model Overview
+
+The model can be expressed mathematically using the logistic function:
+
+P(Y=1 | X) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2)}}
+
+Where:
+
+	•	 P(Y=1 | X)  represents the probability of Trump winning (i.e.,  Y=1 ).
+	•	 \beta_0  is the intercept term.
+	•	 \beta_1  and  \beta_2  are the coefficients for the predictor variables, which include:
+	•	 X_1 : The average polling percentage for Trump ( \text{trump\_pct} ).
+	•	 X_2 : The average polling percentage for Harris ( \text{harris\_pct} ).
+
+Data and Features
+
+The dataset used in this model is sourced from Project 538, focusing on high-quality polling data. Key features include:
+
+	•	trump_pct: The percentage of respondents supporting Trump.
+	•	harris_pct: The percentage of respondents supporting Harris.
+	•	sample_size: The size of each poll, which influences the weight of that poll in our analysis.
+	•	end_date: The date when the poll began, which is used to calculate recency.
+
+Weight Calculation
+
+To ensure that more recent and larger polls have a higher influence on the model, we calculate a weight for each poll as follows:
+
+	1.	Recency Weight:
+
+\text{recency\_weight} = \max(\text{current\_date} - \text{start\_date}) - (\text{current\_date} - \text{start\_date})
+
+This transformation ensures that more recent polls receive a larger weight.
+	2.	Total Weight:
+
+\text{weight} = \text{sample\_size} \times \text{recency\_weight}
+
+The weighted polling percentages for Trump and Harris are then calculated for each state, allowing for a more nuanced aggregation of voter sentiment.
+
+Logistic Regression Implementation
+
+The logistic regression model is implemented using the LogisticRegression class from the sklearn library. The predictor variables are defined as:
+
+	•	 X  = [trump_pct, harris_pct]
+
+The binary outcome variable is defined as follows:
+
+	•	 y = 1  if  \text{trump\_pct} > \text{harris\_pct}  (Trump is predicted to win).
+	•	 y = 0  otherwise.
+
+Model Validation and Assumptions
+
+The assumptions underlying this model include:
+
+	•	Linearity: The log-odds of the outcome variable are linearly related to the predictor variables.
+	•	Independence: Observations are independent of one another.
+
+Limitations
+
+	1.	Polling Bias: The model is contingent upon the quality of the polling data.
+Polls with lower numeric grades or transparency scores are excluded, potentially introducing bias if high-quality polls are not representative of the entire electorate.
+	2.	Dynamic Voter Sentiment: Voter preferences can change rapidly, particularly in the lead-up to the election, and the model may not fully capture these shifts if not updated frequently.
+
+Alternative Models Considered
+
+Alternative models, such as Bayesian logistic regression, were considered.
+While Bayesian models allow for the incorporation of prior distributions and might provide additional insights through credible intervals, they require more complex implementation and careful selection of priors.
+Given the nature of the data and the objective of clear interpretability, the logistic regression model was chosen for its straightforward application and interpretability.
+
+Conclusion
+
+The logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data.
+By focusing on key features such as polling percentages and sample sizes, the model captures the competitive dynamics between candidates and offers valuable insights into potential election outcomes.
+Future iterations of this model can incorporate real-time data updates to enhance predictive accuracy as the election approaches.
+ # 5 Results {#sec-results}
+ # 6 Discussion {#sec-discussion}
+
+ # Apendix {-}
+
 # References

### 2024-10-23 paper/paper.qmd
update paper.qmd with discussion section

@@
 Future iterations of this model can incorporate real-time data updates to enhance predictive accuracy as the election approaches.
- # 5 Results {#sec-results}
- # 6 Discussion {#sec-discussion}
+
+ # 4 Results {#sec-results}
+The results of our analysis are as follows.
+We have Donald Trump predicted to win Arizona, Georgia and Michigan for a total of 42 of the 93 swing state votes.
+While Kamala Harris is projected to win Nevada, North Carolina, Pennsylvania and Wisconsin for the remaining 51 swing state votes.
+This results in a win for the democrats and the first female US president!
+
+# 5 Discussion {#sec-discussion}
+
+## What Is Done in This Paper?
+
+In this paper, we create a logistic regression model to predict the outcome of the 2024 U.S.
+Presidential Election, focusing on state-level polling data to estimate the probability of Donald Trump winning against Kamala Harris.
+By aggregating recent and reliable polls for swing states and weighting them by sample size and recency, we aim to provide a data-driven analysis of voter preferences.
+Our model is constructed to identify the likelihood of Trump securing a majority in each state, ultimately offering a prediction for the overall election result.
+
+## What Do We Learn About the World?
+
+One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets.
+While the majority of reputable polls suggest that Kamala Harris is favored on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
+This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.
+
+At first glance, one might expect that if polling data were more accurate than betting markets, arbitrage opportunities would emerge, allowing savvy participants to profit from discrepancies.
+However, this does not seem to be happening, which suggests that the markets may be pricing in information that the polls do not capture—perhaps reflecting shifts in voter sentiment, hidden preferences, or systematic biases in polling.
+
+## What Else Do We Learn About the World?
+
+Another important consideration is the possible biases in both polling and betting markets.
+Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favor of Harris.
+Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favor him.
+Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities.
+However, historical data still suggest that polls have been accurate 78% of the time in predicting elections [@FiveThirtyEight], indicating that the reliability of betting markets remains questionable, especially in light of their poor performance during the 2016 election, when Trump’s odds were listed at +475 (17%) just before his victory [https://www.oddsshark.com/entertainment/us-presidential-odds-2016-futures].
+
+## Weaknesses of the Model
+
+While our model provides a structured framework for predicting the election, it has several limitations.
+First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals).
+Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve non-linear influences that are difficult to predict.
+
+Moreover, our focus on swing states limits the model’s applicability to the national picture.
+While swing states are crucial to the election outcome, non-swing states could offer additional insights into broader voter trends, and including them in future analyses could enhance the model’s robustness.
+Training the model on a more comprehensive dataset that includes these states and other predictive variables, such as demographic factors or economic indicators, could lead to a more accurate forecast.
+
+## How Should We Proceed in the Future?
+
+Looking ahead, future iterations of this model should incorporate more diverse data sources.
+Expanding the dataset to include polling data from all states, along with betting market information, could improve prediction accuracy.
+Moreover, experimenting with different types of predictive models, such as Bayesian logistic regression or machine learning models like random forests, could offer more sophisticated insights and account for non-linear interactions that our current model might miss.
+
+It would also be valuable to study the interaction between polling data and betting markets more closely, potentially integrating them into a unified prediction model.
+This hybrid approach might help reconcile the discrepancies observed between the two sources and provide a more nuanced understanding of election dynamics.
+Additionally, out-of-sample validation and sensitivity analyses should be conducted to test the robustness of our model and adjust for overfitting.
+
+Ultimately, while our model offers a strong foundation for predicting the 2024 U.S.
+Presidential Election, there is still much to learn.
+By refining the model and incorporating additional data, we can move closer to producing predictions that more accurately reflect voter behavior and election outcomes.
 

### 2024-10-23 paper/paper.qmd
data section update

@@
 # 2.1 Measurement
+
 The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
-This dataset reflects potential outcomes for the 2024 U.S.
+This dataset provides predictions for candidate support for the 2024 U.S.
 Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
 
-It’s important to note that each pollster employs unique sampling techniques and methodologies, which influence their numeric grades and transparency scores.
-To ensure the integrity of our analysis, we selected only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
+The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
+By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
+This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+
+# The dataset
+
+A swing state in the context of a election defined as
+
+It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
+Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
+a good score reflecing low propensity to be impacted by erors and biases tha can occur in survey sampling.
+It also accounts for impacts that might bias a surveys score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
+Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
+By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
+We use this single measurement out of 3 to ensure the integrity of our analysis.
+Selecting only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
 This approach is used so that the data reflects a robust and credible representation of voter preferences.
 
-The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
-By aggregating these individual responses, we are able to derive state-level averages for both candidates.
-This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+In keeping with this aim, we focused our analysis on polls conducted after September.
+Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, by example Biden dropping out as a candidate on July 21st.
+These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
+By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
 

### 2024-10-24 paper/paper.qmd
data section

@@
 
-# The dataset
+# 2.1 The Dataset
 
-A swing state in the context of a election defined as
-
+A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
+States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
+Thus the election result is normally decided by which party wins just a handful of states in the electoral collegate system.
+By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
+Thus our analysis rstricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
 It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
@@
 By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
-
- # 2.1 The Dataset
 

### 2024-10-24 scripts/graph_modelling/key_variabls_plot.py
Some random graphs

@@ -1,0 +1,25 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Load the dataset
+data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")  # Replace with your file path
+
+# Convert the end_date to datetime for plotting
+data['end_date'] = pd.to_datetime(data['end_date'])
+
+# Scatter plot: Poll Score over Time, Colored by State
+plt.figure(figsize=(12, 6))
+states = data['state'].unique()
+
+# Plot each state's data with a different color
+for state in states:
+    subset = data[data['state'] == state]
+    plt.scatter(subset['end_date'], subset['numeric_grade'], label=state, alpha=0.6)
+
+plt.title('poll grade over Time by State')
+plt.xlabel('End Date')
+plt.ylabel('grade Score')
+plt.xticks(rotation=45)
+plt.legend(loc='upper left', bbox_to_anchor=(1, 1), fontsize='small')
+plt.tight_layout()
+plt.show()

### 2024-10-24 scripts/graph_modelling/step_up_by_state.py

@@ -1,0 +1,55 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+# Step 1: Define the Electoral Votes for Each State
+states = ['California', 'Colorado', 'Connecticut', 'Delaware', 'District of Columbia', 'Hawaii', 'Illinois',  'Maryland', 'Massachusetts', 'Minnesota', 'New Hampshire',
+           'New Jersey', 'New Mexico', 'New York',  'Oregon','Washington', 'Arizona', 'Pennsylvania', 'North Carolina', 'Georgia', 'Nevada', 'Michigan', 'Wisconsin',  'Washington', 'Alabama', 'Alaska', 'Arkansas', 'Florida', 'Idaho', 'Indiana', 'Iowa', 'Kansas',  'Kentucky', 'Louisiana',
+            'Maine', 'Mississippi', 'Missouri', 'Montana', 'Nebraska', 'North Dakota',  'Ohio', 'Oklahoma', 'South Carolina', 'South Dakota',
+           'Tennessee', 'Texas', 'Utah', 'West Virginia', 'Wyoming']
+
+
+
+votes = [54, 10, 7, 3, 3, 4, 19, 10, 11, 10, 4, 14, 5, 28, 8, 11, 19, 16, 16, 6, 15, 10, 
+ 12, 9, 3, 6, 30, 4, 11, 6, 6, 8, 8, 4, 6, 10, 4, 5, 3, 17, 7, 9, 3, 11, 40, 6, 3, 10, 3]
+
+# Step 2: Compute the Cumulative Electoral Votes
+cumulative_votes = np.cumsum(votes)
+
+threshold_index = np.argmax(cumulative_votes >= 225)  # First point where cumulative votes reach 270
+x = np.arange(len(cumulative_votes))  # X-axis values (state index)
+
+# Step 3: Plot the Step Function
+plt.figure(figsize=(18, 8))
+plt.step(range(len(cumulative_votes)), cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
+
+# Step 4: Customize the Plot
+plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
+plt.yticks(np.arange(0, 550, 50))
+
+# Fill the area under the curve with blue where Democrats are leading
+plt.fill_between(x[:threshold_index + 1], cumulative_votes[:threshold_index + 1], 
+                 step='post', color='blue', alpha=0.4)
+
+# Add a horizontal line at 270 votes (threshold to win)
[…]

### 2024-10-24 paper/paper.qmd
spellchck chapter 2

@@
 Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
 
@@
 By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
-This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+This transformation of general voter sentiment into specific data points enables us to analyse the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
 
@@
 States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
-Thus the election result is normally decided by which party wins just a handful of states in the electoral collegate system.
+Thus the election result is normally decided by which party wins just a handful of states in the electoral college system.
 By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
-Thus our analysis rstricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
+Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
 It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
 Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
-a good score reflecing low propensity to be impacted by erors and biases tha can occur in survey sampling.
-It also accounts for impacts that might bias a surveys score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
+a good score reflecting low propensity to be impacted by errors and biases that can occur in survey sampling.
+It also accounts for impacts that might bias a survey's score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
 Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
@@
 In keeping with this aim, we focused our analysis on polls conducted after September.
-Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, by example Biden dropping out as a candidate on July 21st.
+Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, for example Biden dropping out as a candidate on July 21st.
 These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.

### 2024-10-24 paper/paper.qmd
update paper.qmd and add the pdf

--- 2024-10-23 paper/paper.qmd
@@
 # 2.1 Measurement 
+The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
+This dataset reflects potential outcomes for the 2024 U.S.
+Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
+Each entry in the dataset corresponds to a specific poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
 
-The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
-This dataset provides predictions for candidate support for the 2024 U.S.
-Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+It’s important to note that each pollster employs unique sampling techniques and methodologies, which influence their numeric grades and transparency scores.
+To ensure the integrity of our analysis, we selected only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
+This approach is used so that the data reflects a robust and credible representation of voter preferences.
 
 The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
-By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
+By aggregating these individual responses, we are able to derive state-level averages for both candidates.
 This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
-
-# The dataset
-
-A swing state in the context of a election defined as
-
-It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
-Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
-a good score reflecing low propensity to be impacted by erors and biases tha can occur in survey sampling.
-It also accounts for impacts that might bias a surveys score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
-Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
-By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
-We use this single measurement out of 3 to ensure the integrity of our analysis.
-Selecting only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
-This approach is used so that the data reflects a robust and credible representation of voter preferences.
-
-In keeping with this aim, we focused our analysis on polls conducted after September.
-Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, by example Biden dropping out as a candidate on July 21st.
-These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
-By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
 
@@
 
-P(Y=1 | X) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2)}}
+$P(Y=1 | X) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \beta_2 X_2)}}$
 
@@
 
-	•	 P(Y=1 | X)  represents the probability of Trump winning (i.e.,  Y=1 ).
-	•	 \beta_0  is the intercept term.
-	•	 \beta_1  and  \beta_2  are the coefficients for the predictor variables, which include:
+•	 $P(Y=1 | X)$  represents the probability of Trump winning (i.e.,  Y=1 ).
+•	 $\beta_0$  is the intercept term.
+•	 $\beta_1$  and  $\beta_2$  are the coefficients for the predictor variables, which include:
 •	 X_1 : The average polling percentage for Trump ( \text{trump\_pct} ).
@@
 
-	1.	Recency Weight:
+Recency Weight:
 
-\text{recency\_weight} = \max(\text{current\_date} - \text{start\_date}) - (\text{current\_date} - \text{start\_date})
+$\text{recency\_weight} = \max(\text{current\_date} - \text{start\_date}) - (\text{current\_date} - \text{start\_date})$
 
 This transformation ensures that more recent polls receive a larger weight.
-	2.	Total Weight:
 
-\text{weight} = \text{sample\_size} \times \text{recency\_weight}
+Total Weight:
+
+$\text{weight} = \text{sample\_size} \times \text{recency\_weight}$
 
@@
 
-	•	 y = 1  if  \text{trump\_pct} > \text{harris\_pct}  (Trump is predicted to win).
-	•	 y = 0  otherwise.
+$y = 1$  if  $\text{trump\_pct} > \text{harris\_pct}$  (Trump is predicted to win)
+y = 0  otherwise.
 
@@
 
- # Apendix {-}
+# Apendix
 
+Appendix: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting
+
+In this revised approach, we will allocate the $100,000 budget not only for data collection and survey methods but also to incentivize participants to bet on the outcome of the election.
+This adds a monetary incentive for respondents to make accurate predictions, aligning their responses with their genuine expectations and potentially enhancing the predictive value of the data.
+Below is an outline of the methodology that integrates incentivized betting, sampling, recruitment, data validation, poll aggregation, and other features.
+
+1. Sampling Approach
+
+We will use stratified random sampling to capture a representative sample of the electorate across key demographics and regions.
+This ensures that our poll and betting pool reflects the diversity of the voting population, with emphasis on:
+
+Demographic groups: Stratified by race, age, gender, education, and income.
+Geographic regions: Oversampling key battleground states (e.g., Pennsylvania, Michigan, Wisconsin, Florida) to ensure robust state-level data while maintaining a representative national sample.
+Voter history: Stratified by past voter turnout, targeting likely voters, less frequent voters, and even non-voters to capture a full spectrum of voter engagement.
+
+The goal is to survey 5,000 respondents, with a significant proportion from battleground states.
+By incentivizing respondents with the betting opportunity, the aim is to foster thoughtful responses, as they will have a monetary stake in their predictions.
+
+2. Recruitment of Respondents
+
+We will use a mixed-mode recruitment strategy, relying on both online and telephone methods to reach a broad audience:
+
+Online recruitment: Participants will be recruited via digital platforms such as social media, news websites, and online political forums.
+Advertisements will emphasize the chance to bet on election outcomes and win money based on the accuracy of their predictions.
+Telephone recruitment: For older demographics and those less reachable online, telephone interviews will be conducted using voter registration rolls and commercial databases.
+
+Incentives: The primary incentive is the betting pool, with $50,000 of the budget allocated to a reward fund for accurate predictions.
+Respondents will have the opportunity to place a bet (e.g., $10) on the candidate they believe will win, and those who predict correctly will share in the prize pool.
+This financial incentive should increase engagement and encourage respondents to think carefully about their answers.
+
+3. Survey and Betting Design
+
+The survey will gather standard polling data, but with the addition of a betting component:
+
+Candidate preference: “If the 2024 U.S. Presidential Election were held today, would you vote for?”
+•	Kamala Harris (D)
+•	Donald Trump (R)
+•	Other/Third-party candidate
+•	Undecided
+Betting decision: After stating their candidate preference, respondents will be asked: “As a part of this survey you have been allotted $10 to bet on who you believe is most likely to win the election given the current odds.
+If the candidate of your choosing wins you will recieve a payout equal to the live odds.
+Which candidate would you like to place a bet on?” They will be given $10 to place on the candidate they believe is most likely to win.
+
+Voter behavior and demographics: Questions about voter intention, past voting behavior, and demographic details will follow, ensuring the data can be properly stratified and analyzed.
+
+Betting market engagement: “Have you participated in any betting markets (e.g., Polymarket, PredictIt) related to the election?” This will help analyze how participants in betting markets differ from typical poll respondents.
+
+Election forecasting confidence: “On a scale of 1-10, how confident are you in your prediction of who will win the election?”
+
+4. Data Validation
+
+To ensure the quality of the data, several validation measures will be implemented:
+
+•	Consistency checks: Responses will be monitored for inconsistencies 
+•	Verification: Voter registration data will be cross-referenced to verify the eligibility of respondents.
+•	Weighting: Data will be weighted to reflect the broader electorate using Census data on demographics and voting patterns.
+
+5. Poll Aggregation and Modeling
+
+After collecting the data, we will aggregate our polling results alongside other reliable polls and incorporate betting market data.
+This will allow us to adjust for discrepancies between public opinion polling and market-based predictions.
+
+•	Poll aggregation: Our results will be integrated with other national and state-level polls from reputable sources, such as Gallup and YouGov.
+Polls will be weighted based on sample size, recency, and historical accuracy.
+•	Betting market integration: We will compare the betting behaviors within our sample to existing market data from platforms like PredictIt and Polymarket.
+This comparison will help adjust for potential biases in polling, as betting markets can reveal hidden preferences or voter reluctance to openly state their choices.
+•	Modeling: We will use Bayesian hierarchical models to predict the election outcome at both the national and state levels.
+The model will integrate polling data, betting odds, and historical election trends.
+By comparing how betting and polling data align or diverge, we will refine the forecast, accounting for last-minute voter shifts or surprise outcomes.
+
+6. Poll-Betting Market Discrepancy
+
+A major component of this methodology is addressing the discrepancies between polling and betting markets.
+Our hypothesis is that participants are more likely to provide accurate predictions when they have a financial stake in the outcome.
+
+•	Betting behavior analysis: We will analyze how betting behavior correlates with stated voter preferences, identifying whether betting markets capture sentiment that traditional polling does not.
+•	Incentive impact: By comparing the results of those who opted to bet versus those who did not, we can determine whether financial incentives lead to more accurate election predictions.
+
+7. Budget Breakdown
+
+The $100,000 budget will be allocated as follows:
+
+•	Respondent recruitment: $10,000
+•	Survey administration (online and telephone): $20,000
+•	Betting pool incentives: $50,000 (for those who accurately predict the election outcome)
+•	Data validation and analysis: $10,000
+•	Poll aggregation and modeling: $10,000
+
+Conclusion
+
+This methodology uses a novel approach to incentivize accurate predictions by allowing respondents to bet on the election outcome.
+By combining traditional polling techniques with betting incentives, we aim to close the gap between polling data and betting market predictions, potentially uncovering hidden voter behaviors or biases that standard polling methods may miss.
+
+The survey can be found here:
+https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0
 # References

### 2024-10-24 paper/paper.qmd

@@
 # 2.1 Measurement
+
 The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
-This dataset reflects potential outcomes for the 2024 U.S.
+This dataset provides predictions for candidate support for the 2024 U.S.
 Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
 
-It’s important to note that each pollster employs unique sampling techniques and methodologies, which influence their numeric grades and transparency scores.
-To ensure the integrity of our analysis, we selected only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
+The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
+By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
+This transformation of general voter sentiment into specific data points enables us to analyse the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+
+# 2.1 The Dataset
+
+A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
+States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
+Thus the election result is normally decided by which party wins just a handful of states in the electoral college system.
+By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
+Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
+It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
+Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
+a good score reflecting low propensity to be impacted by errors and biases that can occur in survey sampling.
+It also accounts for impacts that might bias a survey's score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
+Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
+By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
+We use this single measurement out of 3 to ensure the integrity of our analysis.
+Selecting only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
 This approach is used so that the data reflects a robust and credible representation of voter preferences.
 
-The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
-By aggregating these individual responses, we are able to derive state-level averages for both candidates.
-This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
-
- # 2.1 The Dataset
+In keeping with this aim, we focused our analysis on polls conducted after September.
+Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, for example Biden dropping out as a candidate on July 21st.
+These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
+By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
 

### 2024-10-24 scripts/00-simulate_data.py
Update the preamble for the scripts

@@ -1,25 +1,15 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  # state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates US presidential election polling data
+# Author: Alexander Guarasci & Jacob Gilbert
+# Date: November 4, 2024
+# Contact: alexander.guarasci@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - `polars` must be installed (pip install polars)
+  # - `pandas` must be installed (pip install polars)
   # - `numpy` must be installed (pip install numpy)
 
 
-#### Workspace setup ####
-#### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+
 
 
 import pandas as pd

### 2024-10-24 scripts/01-test_simulated_data.py

@@ -1,3 +1,13 @@
+#### Preamble ####
+# Purpose: Tests the simulated US election polling data
+# Author: Alexander Guarasci & Jacob Gilbert
+# Date: November 4, 2024
+# Contact: alexander.guarasci@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - `pandas` must be installed (pip install polars)
+  # - `unittest` must be installed (pip install unittest)
+
 import pandas as pd
 import unittest
 

### 2024-10-25 scripts/03-clean_data.py
move from csv format to parquet

@@ -11,6 +11,7 @@
 #### Workspace setup ####
 import pandas as pd
 
+
 raw_df = pd.read_csv("data/raw_data/president_polls.csv")
 
 selected_columns = [
@@ -41,8 +42,8 @@ df = raw_df[selected_columns]
 df = df.loc[df['numeric_grade'] >= 2.5]
 df = df[df['candidate_name'].isin(['Donald Trump', 'Kamala Harris'])]
 #filter the recent 
-df['poll_end_date'] = pd.to_datetime(df['end_date'], format='%m/%d/%y')
-df = df.loc[df["poll_end_date"] > "2024-09-01"]
+df['end_date'] = pd.to_datetime(df['end_date'], format='%m/%d/%y')
+df = df.loc[df["end_date"] > "2024-09-01"]
 
 df_swing = df.loc[df["state"].isin(["Arizona","Pennsylvania", "North Carolina", "Georgia", "Nevada", "Michigan", "Wisconsin"])]
 
@@ -62,7 +63,7 @@ print(df_swing['unique_id'].nunique())
 
 print(df_swing["numeric_grade"].min())
 # save raw swing stat data
-df_swing.to_csv("data/02-analysis_data/swing_state_polls.csv")
+df_swing.to_parquet("data/02-analysis_data/swing_state_polls.csv")
 
 #creating data to be analysed (grouping by )
 df_trump = df_swing[df_swing['candidate_name'] == "Donald Trump"]
@@ -85,4 +86,4 @@ df_by_poll = pd.merge(df_trump[['unique_id',
     df_harris[['unique_id', 'pct']], on='unique_id', how='inner')
 df_merged = df_by_poll.rename(columns={'pct_x': 'trump_pct', 'pct_y': 'harris_pct'}, inplace=True)
 
-df_by_poll.to_csv("data/02-analysis_data/merged_swing_state_data.csv")
+df_by_poll.to_parquet("data/02-analysis_data/merged_swing_state_data.parquet")

### 2024-10-25 scripts/04-test_analysis_data.py

@@ -12,7 +12,7 @@
 import pandas as pd
 
 # Load the dataset
-df = pd.read_csv('data/02-analysis_data/merged_swing_state_data.csv')
+df = pd.read_csv('data/02-analysis_data/merged_swing_state_data.parquet')
 
 def test_trump_harris_pct_sum():
     """

### 2024-10-25 scripts/xandermodel.py

@@ -2,7 +2,7 @@ import pandas as pd
 from sklearn.linear_model import LogisticRegression
 
 # Load the cleaned polling data
-clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")
+clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.parquet")
 
 # Create a weight for each poll based on recency and sample size
 # For simplicity, let's assume that the weight is based on the sample size directly

### 2024-10-25 scripts/graph_modelling/key_variabls_plot.py
Some graphs and data section first draft

@@ -2,7 +2,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 
 # Load the dataset
-data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")  # Replace with your file path
+data = pd.read_csv("data/02-analysis_data/swing_state_polls.csv")  # Replace with your file path
 
 # Convert the end_date to datetime for plotting
 data['end_date'] = pd.to_datetime(data['end_date'])
@@ -14,7 +14,7 @@ states = data['state'].unique()
 # Plot each state's data with a different color
 for state in states:
     subset = data[data['state'] == state]
-    plt.scatter(subset['end_date'], subset['numeric_grade'], label=state, alpha=0.6)
+    plt.scatter(subset['end_date'], subset['pct'], label=state, alpha=0.6)
 
 plt.title('poll grade over Time by State')
 plt.xlabel('End Date')

### 2024-10-25 scripts/graph_modelling/polls_over_time.py

@@ -1,0 +1,22 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Load your data (replace with your actual data file path)
+data = pd.read_parquet('data/02-analysis_data/merged_swing_state_data.parquet')
+
+# Ensure date column is in datetime format
+data['end_date'] = pd.to_datetime(data['end_date'])
+
+# Plot the histogram of poll quantity by end date
+plt.figure(figsize=(12, 6))
+plt.hist(data['end_date'], bins=20, edgecolor='black', alpha=0.7)
+
+# Adding labels and title
+plt.xlabel('End Date of Poll')
+plt.ylabel('Number of Polls')
+plt.title('Histogram of Poll Quantity by End Date')
+plt.xticks(rotation=45)
+
+# Show the plot
+plt.tight_layout()
+plt.show()

### 2024-10-25 paper/paper.qmd

@@
 ---
+
+```{python}
+#| echo: false
+import pandas as pd
+import matplotlib.pyplot as plt
+import jupyter
+from tabulate import tabulate
+```
 
@@
 
+```{python}
+#| echo: false
+
+# Load the data (replace with your CSV file path)
+df = pd.read_csv('../data/02-analysis_data/merged_swing_state_data.csv')
+
+# Group by state and calculate summary statistics
+summary_stats = df.groupby('state').agg(
+    numeric_grade_mean=('numeric_grade', 'mean'),
+    numeric_grade_std=('numeric_grade', 'std'),
+    numeric_grade_min=('numeric_grade', 'min'),
+    numeric_grade_max=('numeric_grade', 'max'),
+    numeric_grade_count=('numeric_grade', 'count'),
+    sample_size_mean=('sample_size', 'mean'),
+    sample_size_std=('sample_size', 'std'),
+    sample_size_min=('sample_size', 'min'),
+    sample_size_max=('sample_size', 'max'),
+    sample_size_count=('sample_size', 'count')
+).reset_index()
+
+# Reshape the summary table to long format
+summary_long = pd.melt(
+    summary_stats,
+    id_vars='state',
+    value_vars=[
+        'numeric_grade_mean', 'numeric_grade_std', 'numeric_grade_min', 'numeric_grade_max', 'numeric_grade_count',
+        'sample_size_mean', 'sample_size_std', 'sample_size_min', 'sample_size_max', 'sample_size_count'
+    ],
+    var_name='variable_stat', value_name='value'
+)
+
+# Split 'variable_stat' into 'variable' and 'stat'
+summary_long[['variable', 'stat']] = summary_long['variable_stat'].str.rsplit('_', n=1, expand=True)
+
+# Drop the original column
+summary_long.drop('variable_stat', axis=1, inplace=True)
+
+# Pivot the table to get desired output
+summary_pivot = summary_long.pivot_table(
+    index=['state', 'variable'], columns='stat', values='value', aggfunc='first'
+).reset_index()
+
+# Sort for better readability
+summary_pivot = summary_pivot.sort_values(by=['state', 'variable'])
+
+summary_pivot[['mean', 'std', 'min', 'max']] = summary_pivot[['mean', 'std', 'min', 'max']].round(1)
+
+# Create a list of rows for tabulate (merging state rows visually)
+rows = []
+last_state = None
+for _, row in summary_pivot.iterrows():
+    state = row['state']
+    variable = row['variable']
+    mean, std, min_, max_, count = row['mean'], row['std'], row['min'], row['max'], row['count']
+
+    # Only show state name on the first row of each group
+    if state == last_state:
+        state = ''
+    else:
+        last_state = state
+
+    rows.append([state, variable, mean, std, min_, max_, count])
+
+# Print the table using tabulate
+print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='fancy_grid'))
+```
+
  # 2.2 Variables of Interest
+In our logistic regression model predicting the 2024 U.S.
+Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data.
+The percentage of respondant favouring a candidate become the key variables for prediction.
+Within the collection of polls in our sample the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump , and `harris_pct` respectively.
+Polls ar taking towards the build up of an election and thus ar taken by repondants at different times.
+Graph  x shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
+Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
+```{python}
+#| echo: false
+
+# Load your data (replace with your actual data file path)
+data = pd.read_parquet('../data/02-analysis_data/merged_swing_state_data.parquet')
+
+# Ensure date column is in datetime format
+data['end_date'] = pd.to_datetime(data['end_date'])
+
+# Plot the histogram of poll quantity by end date
+plt.figure(figsize=(12, 6))
+plt.hist(data['end_date'], bins=20, edgecolor='black', alpha=0.7)
+
+# Adding labels and title
+plt.xlabel('End Date of Poll')
+plt.ylabel('Number of Polls')
+plt.title('Histogram of Poll Quantity by End Date')
+plt.xticks(rotation=45)
+
+# Show the plot
+plt.tight_layout()
+plt.show()
+
+```
+
+Another aspects of the polls we wish to consider is the population size of each.
+Larger polls result in a stronger reduction in randomness in the response.
+The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers.
+In table x we can see the variety within the states polls.
+For use in the model the percentages in these polls is weighted for recenecy and size of the population.
 
@@
 # Apendix
+(include some detail on how we are getting rid of biass. and focus on how we are spcfically uing stratifid and clustr sampling. how we will reach ou. how we will gt rsponce.)
 
@@
 https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0
+
+# appndix B 
+say what kind of ampling. 
+reference rsearch on wha the sampling mthod is
+
 # References

### 2024-10-25 paper/paper.qmd
Step up graph for non swing states

@@
 Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
+In figure 3 we see the ammount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
+
+```{python}
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Step 1: Define the Electoral Votes for Each State
+states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois",
+"Maine", "Maryland", "Massachusetts", "Minnesota", "New Jersey",
+"New York", "Oregon", "Rhode Island", "Vermon", "Washington",
+"District of Columbia", "New Mexico", "New Hampshire", "Colorado", "Virginia",
+"Arizona",
+"Georgia",
+"Michigan",
+"Nevada",
+"North Carolina",
+"Pennsylvania",
+"Wisconsin",
+"Alabama", "Alaska", "Arkansas", "Florida", "Idaho",
+"Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana",
+"Mississippi", "Missouri", "Montana", "Nebraska", "North Dakota",
+"Ohio", "Oklahoma", "South Carolina", "South Dakota", "Tennessee",
+"Texas", "Utah", "West Virginia", "Wyoming"]
+print(len(states))
+votes = [54, 7 , 2, 4, 19, 4, 10, 11, 10, 14, 28, 8, 4, 3, 12, 3, 5, 4, 10, 13,
+         11,16,15,6,16,19,10,
+         9,3,6,30,4,11,6,6,8,8,6,10,4,5,3,17,7,9,3,11,40,6,4,3]
+
+print(len(votes))
+# Step 2: Compute the Cumulative Electoral Votes
+cumulative_votes = np.cumsum(votes)
+
+# Compute Cumulative Votes in Reverse Order for Republicans
+cumulative_votes_reversed = np.cumsum(votes[::-1])
+
+# X-axis values (state index)
+x = np.arange(len(cumulative_votes))
+
+# Step 3: Plot the Step Function for Democrats (left to right)
+plt.figure(figsize=(18, 8))
+plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
+
+# Reverse the x-axis for the Republican step function (right to left)
+x_reversed = np.flip(x)
+plt.step(x_reversed, cumulative_votes_reversed, where='post', color='red', linewidth=2, marker='o')
+
+# Step 4: Customize the Plot
+plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
+plt.yticks(np.arange(0, 550, 50))
+
+# Fill the area under the curve with blue for Democrats (left to right)
+threshold_index = np.argmax(cumulative_votes >= 227)  # First point where cumulative votes reach 270
+plt.fill_between(x[:threshold_index + 1], cumulative_votes[:threshold_index + 1], 
+                 step='post', color='blue', alpha=0.4)
+
+# Fill the area under the curve with red for Republicans (right to left)
+threshold_index = np.argmax(cumulative_votes >= 219)
+plt.fill_between(x_reversed[:threshold_index + 5], cumulative_votes_reversed[:threshold_index + 5], step='post', color='red', alpha=0.4)
+
+# Add a horizontal line at 270 votes (threshold to win)
+plt.axhline(270, color='purple', linestyle='--', linewidth=2)
+plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
+
+# Add lines for current Democratic and Republican thresholds
+plt.axhline(225, color='blue', linestyle='--', linewidth=1.5)
+plt.text(210, 220, '225 votes', horizontalalignment='center', color='blue', fontsize=12)
+
+# Add Labels and Title
+plt.xlabel('Electoral College Markets (electoral votes)', fontsize=12)
+plt.ylabel('Total Electoral Votes', fontsize=12)
+plt.title('Step Function of Total Electoral Votes\nwith States Ordered by Vote Share', fontsize=15)
+
+# Add Grid for Better Visibility
+plt.grid(True, linestyle='--', alpha=0.5)
+
+# Adjust Layout for Better Fit
+plt.tight_layout()
+
+# Step 5: Show the Plot
+plt.show()
+
+```
+
 It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.

### 2024-10-25 scripts/graph_modelling/step_up_by_state.py

@@ -2,17 +2,27 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 # Step 1: Define the Electoral Votes for Each State
-states = ['California', 'Colorado', 'Connecticut', 'Delaware', 'District of Columbia', 'Hawaii', 'Illinois', 
-          'Maryland', 'Massachusetts', 'Minnesota', 'New Hampshire', 'New Jersey', 'New Mexico', 'New York',  
-          'Oregon', 'Washington', 'Arizona', 'Pennsylvania', 'North Carolina', 'Georgia', 'Nevada', 'Michigan', 
-          'Wisconsin', 'Washington', 'Alabama', 'Alaska', 'Arkansas', 'Florida', 'Idaho', 'Indiana', 'Iowa', 
-          'Kansas', 'Kentucky', 'Louisiana', 'Maine', 'Mississippi', 'Missouri', 'Montana', 'Nebraska', 
-          'North Dakota', 'Ohio', 'Oklahoma', 'South Carolina', 'South Dakota', 'Tennessee', 'Texas', 'Utah', 
-          'West Virginia', 'Wyoming']
+states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois",
+"Maine", "Maryland", "Massachusetts", "Minnesota", "New Jersey",
+"New York", "Oregon", "Rhode Island", "Vermon", "Washington",
+"District of Columbia", "New Mexico", "New Hampshire", "Colorado", "Virginia",
+"Arizona",
+"Georgia",
+"Michigan",
+"Nevada",
+"North Carolina",
+"Pennsylvania",
+"Wisconsin",
+"Alabama", "Alaska", "Arkansas", "Florida", "Idaho",
+"Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana",
+"Mississippi", "Missouri", "Montana", "Nebraska", "North Dakota",
+"Ohio", "Oklahoma", "South Carolina", "South Dakota", "Tennessee",
+"Texas", "Utah", "West Virginia", "Wyoming"]
 print(len(states))
-votes = [54, 10, 7, 3, 3, 4, 19, 10, 11, 10, 4, 14, 5, 28, 8, 11, 19, 16, 16, 6, 15, 10, 
[…]

### 2024-10-25 scripts/xandermodel.py
Vote pct mean table

@@ -2,7 +2,7 @@ import pandas as pd
 from sklearn.linear_model import LogisticRegression
 
 # Load the cleaned polling data
-clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.parquet")
+clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")
 
 # Create a weight for each poll based on recency and sample size
 # For simplicity, let's assume that the weight is based on the sample size directly

### 2024-10-25 scripts/xandermodel.py
Adding stds to regression results

@@ -10,24 +10,28 @@ clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv
 # You can customize the recency weight according to your requirements
 current_date = pd.to_datetime('today')
 clean_poll_data['start_date'] = pd.to_datetime(clean_poll_data['start_date'])
-clean_poll_data['recency_weight'] = (current_date - clean_poll_data['start_date']).dt.days
 
-# Normalize the recency weight (larger weight for more recent polls)
+# Calculate recency weight: newer polls have larger weights
+clean_poll_data['recency_weight'] = (current_date - clean_poll_data['start_date']).dt.days
 clean_poll_data['recency_weight'] = clean_poll_data['recency_weight'].max() - clean_poll_data['recency_weight']
+
+# Calculate the overall weight by multiplying sample size with recency weight
 clean_poll_data['weight'] = clean_poll_data['sample_size'] * clean_poll_data['recency_weight']
 
 # Aggregate polling percentages for Trump and Harris by state, weighted by sample size and recency
 state_average = clean_poll_data.groupby('state').agg(
-    trump_pct=('trump_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
-    harris_pct=('harris_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum())
+    trump_pct_mean=('trump_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
+    harris_pct_mean=('harris_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
+    trump_pct_sd=('trump_pct', 'std'),
+    harris_pct_sd=('harris_pct', 'std')
 ).reset_index()
 
 # Calculate the probabilities of Trump winning based on the weighted percentages
-state_average['trump_win_prob'] = state_average['trump_pct'] / (state_average['trump_pct'] + state_average['harris_pct'])
+state_average['trump_win_prob'] = state_average['trump_pct_mean'] / (state_average['trump_pct_mean'] + state_average['harris_pct_mean'])
 
 # Prepare the logistic regression model
-X = state_average[['trump_pct', 'harris_pct']]
[…]

### 2024-10-25 paper/paper.qmd
Update the model to include binary, made typo fixes to model section of paper

--- 2024-10-24 paper/paper.qmd
@@
 # 2.1 Measurement 
+The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
+This dataset reflects potential outcomes for the 2024 U.S.
+Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
+Each entry in the dataset corresponds to a specific poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
 
-The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
-This dataset provides predictions for candidate support for the 2024 U.S.
-Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+It’s important to note that each pollster employs unique sampling techniques and methodologies, which influence their numeric grades and transparency scores.
+To ensure the integrity of our analysis, we selected only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
+This approach is used so that the data reflects a robust and credible representation of voter preferences.
 
 The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
-By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
-This transformation of general voter sentiment into specific data points enables us to analyse the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+By aggregating these individual responses, we are able to derive state-level averages for both candidates.
+This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
 
  # 2.1 The Dataset
-
-A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
-States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
-Thus the election result is normally decided by which party wins just a handful of states in the electoral college system.
-By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
-Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
-It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
-Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
-a good score reflecting low propensity to be impacted by errors and biases that can occur in survey sampling.
-It also accounts for impacts that might bias a survey's score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
-Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
-By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
-We use this single measurement out of 3 to ensure the integrity of our analysis.
-Selecting only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
-This approach is used so that the data reflects a robust and credible representation of voter preferences.
-
-In keeping with this aim, we focused our analysis on polls conducted after September.
-Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, for example Biden dropping out as a candidate on July 21st.
-These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
-By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
 
@@
 	•	sample_size: The size of each poll, which influences the weight of that poll in our analysis.
-	•	end_date: The date when the poll began, which is used to calculate recency.
+	•	start_date: The date when the poll began, which is used to calculate recency.
 
@@
 
-$\text{recency\_weight} = \max(\text{current\_date} - \text{start\_date}) - (\text{current\_date} - \text{start\_date})$
+$\text{recency weight} = \max(\text{current date} - \text{start date}) - (\text{current date} - \text{start date})$
 
@@
 
-$\text{weight} = \text{sample\_size} \times \text{recency\_weight}$
+$\text{weight} = \text{sample size} \times \text{recency weight}$
 

### 2024-10-25 scripts/xandermodel.py

@@ -36,9 +36,19 @@ logistic_model.fit(X, y)
 # Predict probabilities of Trump winning
 state_average['predicted_win_prob'] = logistic_model.predict_proba(X)[:, 1]
 
+
 # Output the results
 final_results = state_average[['state', 'trump_pct', 'harris_pct', 'predicted_win_prob']].copy()
 final_results['predicted_win_prob'] = final_results['predicted_win_prob'] * 100  # Convert to percentage
 final_results.columns = ['State', 'Trump %', 'Harris %', 'Percentage Chance of Trump Winning']
 
 print(final_results)
+
+
+
+
+
+
+
+
+

### 2024-10-25 scripts/00-simulate_data.py
results bar graph

@@ -71,6 +71,4 @@ df = pd.DataFrame(data, columns=[
 print(df.head())
 
 # Save the simulated dataset to a CSV file
-df.to_csv("simulated_polling_data.csv", index=False)
-# Save the simulated dataset to a CSV file
 df.to_csv("data/00-simulated_data/simulated_data.csv", index=False)

### 2024-10-25 scripts/graph_modelling/bar_plot_of_results.py

@@ -1,0 +1,39 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+
+# Sample data
+data = {
+    'State': ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 
+              'Pennsylvania', 'Wisconsin'],
+    'Trump % (Mean)': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
+    'Trump % (SD)': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
+    'Harris % (Mean)': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
+    'Harris % (SD)': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
+    'Percentage Chance of Trump Winning': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
+}
+
+# Create DataFrame
+df = pd.DataFrame(data)
+
+# Calculate the distance from 50%
+df['Distance from 50'] = df['Percentage Chance of Trump Winning'] - 50
+
+# Set up the plot
+fig, ax = plt.subplots(figsize=(10, 6))
+
+# Plot bars: left if below 50, right if above 50
+colors = ['blue' if x < 0 else 'red' for x in df['Distance from 50']]
+ax.barh(df['State'], df['Distance from 50'], color=colors, align='center')
+
+# Draw a vertical line at 0 (representing 50%)
+ax.axvline(0, color='black', linewidth=1)
+
+# Set labels and title
+ax.set_xlabel('Distance from 50% (Trump Winning Probability)')
+ax.set_title('Centered Bar Chart of Trump’s Winning Probability by State')
+
+# Adjust layout
+plt.tight_layout()
+
+# Show plot
+plt.show()

### 2024-10-25 scripts/xandermodel.py

@@ -59,3 +59,4 @@ final_results = final_results.round(2)
 
 # Print the final table
 print(final_results)
+final_results.to_csv("final_results.csv")

### 2024-10-25 scripts/xandermodel1.py
model section editing

@@ -1,64 +1,44 @@
-# Import necessary libraries
 import pandas as pd
 from sklearn.linear_model import LogisticRegression
-from sklearn.preprocessing import StandardScaler
-import joblib
 
 # Load the cleaned polling data
-data_path = "data/02-analysis_data/merged_swing_state_data.csv"
-clean_poll_data = pd.read_csv(data_path)
+clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")
 
-# Check if necessary columns exist
-required_columns = ['trump_pct', 'harris_pct', 'end_date', 'sample_size', 'state']
-for column in required_columns:
-    if column not in clean_poll_data.columns:
-        raise ValueError(f"The required column '{column}' is not present in the dataset.")
+# Create a weight for each poll based on recency and sample size
+current_date = pd.to_datetime('today')
+clean_poll_data['start_date'] = pd.to_datetime(clean_poll_data['start_date'])
+clean_poll_data['recency_weight'] = (current_date - clean_poll_data['start_date']).dt.days
 
-# Convert end_date column to datetime format
-clean_poll_data['end_date'] = pd.to_datetime(clean_poll_data['end_date'])
-
-# Define a weighting function
-def calculate_weight(row):
-    # Weight based on recency (more recent = higher weight)
-    recency_weight = 1 / (pd.Timestamp.now() - row['end_date']).days  # More recent polls have higher weights
-    sample_size_weight = row['sample_size'] / clean_poll_data['sample_size'].max()  # Scale by max sample size
-    return recency_weight * sample_size_weight
-
-# Apply the weighting function to create a new 'weight' column
-clean_poll_data['weight'] = clean_poll_data.apply(calculate_weight, axis=1)
+# Normalize the recency weight (larger weight for more recent polls)
+clean_poll_data['recency_weight'] = clean_poll_data['recency_weight'].max() - clean_poll_data['recency_weight']
+clean_poll_data['weight'] = clean_poll_data['sample_size'] * clean_poll_data['recency_weight']
 
 # Aggregate polling percentages for Trump and Harris by state, weighted by sample size and recency
 state_average = clean_poll_data.groupby('state').agg(
     trump_pct=('trump_pct', lambda x: (x * clean_poll_data.loc[x.index, 'weight']).sum() / clean_poll_data.loc[x.index, 'weight'].sum()),
[…]

### 2024-10-28 paper/paper.qmd
Updated appendix a

--- 2024-10-25 paper/paper.qmd
@@
 
-# Apendix
-(include some detail on how we are getting rid of biass. and focus on how we are spcfically uing stratifid and clustr sampling. how we will reach ou. how we will gt rsponce.)
+# Appendix A: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting
 
-Appendix: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting
+## 1. Sampling Approach
 
-In this revised approach, we will allocate the $100,000 budget not only for data collection and survey methods but also to incentivize participants to bet on the outcome of the election.
-This adds a monetary incentive for respondents to make accurate predictions, aligning their responses with their genuine expectations and potentially enhancing the predictive value of the data.
-Below is an outline of the methodology that integrates incentivized betting, sampling, recruitment, data validation, poll aggregation, and other features.
+To ensure that our sample represents the diversity of the US electorate and minimizes potential biases, we employ stratified and cluster sampling techniques.
+With a budget of $100,000, our goal is to gather data from 5,000 respondents, focusing on key swing states while preserving national representativeness.
+The sample will be stratified by:
 
-1. Sampling Approach
+- **Demographics**: Race, age, gender, education, and income.
+- **Geography**: Emphasis on swing states with representation from urban, suburban, and rural areas.
+- **Voter History**: Including respondents with varying voting patterns, such as frequent voters, occasional voters, and those who are less likely to vote.
 
-We will use stratified random sampling to capture a representative sample of the electorate across key demographics and regions.
-This ensures that our poll and betting pool reflects the diversity of the voting population, with emphasis on:
+To address geographic and political biases, cluster sampling will target diverse regions within each state (urban, suburban, and rural areas), capturing the variation in political leanings.
+Additionally, we will soft-launch the survey in order to catch potential issues.
 
-Demographic groups: Stratified by race, age, gender, education, and income.
-Geographic regions: Oversampling key battleground states (e.g., Pennsylvania, Michigan, Wisconsin, Florida) to ensure robust state-level data while maintaining a representative national sample.
-Voter history: Stratified by past voter turnout, targeting likely voters, less frequent voters, and even non-voters to capture a full spectrum of voter engagement.
+## 2. Recruitment of Respondents
 
-The goal is to survey 5,000 respondents, with a significant proportion from battleground states.
-By incentivizing respondents with the betting opportunity, the aim is to foster thoughtful responses, as they will have a monetary stake in their predictions.
+Respondents will be recruited through a multi-channel outreach strategy that combines online and telephone efforts to ensure a representative sample:
 
-2. Recruitment of Respondents
+- **Online Recruitment**: Targeted ads on social media, political forums, and news websites will attract a wide audience, highlighting the unique opportunity to place a small bet on the election outcome.
+- **Telephone Recruitment**: To capture older demographics and those less reachable online, we will conduct phone outreach using commercial databases and voter registration information.
 
-We will use a mixed-mode recruitment strategy, relying on both online and telephone methods to reach a broad audience:
+### Incentives for Engagement
+A betting pool is a central part of this methodology, with $50,000 allocated as rewards for respondents who accurately predict the election outcome.
+Each participant will receive a $10 allowance to bet on their predicted winner (Trump or Harris), with payouts based on live odds, creating a direct incentive for respondents to make thoughtful, informed predictions.
+The purpose of this is twofold, it aims to minmize the attrition of the respondants, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win.
+Ideally, the purpose of this is so that people take into consideration who their friends and family, as well as their community are expecting to win.
 
-Online recruitment: Participants will be recruited via digital platforms such as social media, news websites, and online political forums.
-Advertisements will emphasize the chance to bet on election outcomes and win money based on the accuracy of their predictions.
-Telephone recruitment: For older demographics and those less reachable online, telephone interviews will be conducted using voter registration rolls and commercial databases.
+## 3. Data Validation and Bias Reduction
 
-Incentives: The primary incentive is the betting pool, with $50,000 of the budget allocated to a reward fund for accurate predictions.
-Respondents will have the opportunity to place a bet (e.g., $10) on the candidate they believe will win, and those who predict correctly will share in the prize pool.
-This financial incentive should increase engagement and encourage respondents to think carefully about their answers.
+To further reduce biases, several measures will be implemented for data validation and weighting:
 
-3. Survey and Betting Design
+- **Cross-Referencing Survey Responses**: Survey responses will be cross-referenced with voter registration data to validate eligibility.
+- **Weighting**: Data will be weighted to reflect broader population demographics, adjusting for any imbalances in representation across age, race, gender, and geographic factors.
+This ensures that the results better mirror the entire electorate.
 
-The survey will gather standard polling data, but with the addition of a betting component:
+## 4. Poll Aggregation
 
-Candidate preference: “If the 2024 U.S. Presidential Election were held today, would you vote for?”
-•	Kamala Harris (D)
-•	Donald Trump (R)
-•	Other/Third-party candidate
-•	Undecided
-Betting decision: After stating their candidate preference, respondents will be asked: “As a part of this survey you have been allotted $10 to bet on who you believe is most likely to win the election given the current odds.
-If the candidate of your choosing wins you will recieve a payout equal to the live odds.
-Which candidate would you like to place a bet on?” They will be given $10 to place on the candidate they believe is most likely to win.
+Data collected through this survey will be aggregated with other national and state-level surveys to improve accuracy. The aggregation process will account for:
 
-Voter behavior and demographics: Questions about voter intention, past voting behavior, and demographic details will follow, ensuring the data can be properly stratified and analyzed.
+- **Sample Size and Recency**: More recent polls and those with larger sample sizes will be weighted more heavily.
+- **Pollster Reliability**: Historical poll accuracy and transparency scores will further inform the weight of each poll.
 
-Betting market engagement: “Have you participated in any betting markets (e.g., Polymarket, PredictIt) related to the election?” This will help analyze how participants in betting markets differ from typical poll respondents.
+## 5. Budget Allocation
 
-Election forecasting confidence: “On a scale of 1-10, how confident are you in your prediction of who will win the election?”
+The proposed $100,000 budget will be distributed as follows:
 
-4. Data Validation
+- **Recruitment**: $10,000 for targeting and engaging respondents across multiple platforms.
+- **Survey Administration (Online and Phone)**: $20,000 for both online and phone-based survey collection.
+- **Betting Pool**: $50,000 allocated to reward accurate predictions and enhance response quality.
+- **Data Validation and Analysis**: $10,000 for verification and weighting.
+- **Modeling**: $10,000 for analyzing and forecasting based on aggregated data.
 
-To ensure the quality of the data, several validation measures will be implemented:
+## 6. Question Design and Goal Allignment
+The goal of our survey is to predict who will win the election, this is the research question that guides our survey.
+In order to achieve this, our questions will be isolate the relevant variables by asking only one thing at a time (ceteris paribus).
+We will also use item specific scales over agree-disagree formats to avoid acquiescence bias.
+Furthermore, we will randomize response options in order to mitigate response order bias, and emphasize the surveys anonymity to minimize social desireablity bias.
 
-•	Consistency checks: Responses will be monitored for inconsistencies 
-•	Verification: Voter registration data will be cross-referenced to verify the eligibility of respondents.
-•	Weighting: Data will be weighted to reflect the broader electorate using Census data on demographics and voting patterns.
-
-5. Poll Aggregation and Modeling
-
-After collecting the data, we will aggregate our polling results alongside other reliable polls and incorporate betting market data.
-This will allow us to adjust for discrepancies between public opinion polling and market-based predictions.
-
-•	Poll aggregation: Our results will be integrated with other national and state-level polls from reputable sources, such as Gallup and YouGov.
-Polls will be weighted based on sample size, recency, and historical accuracy.
-•	Betting market integration: We will compare the betting behaviors within our sample to existing market data from platforms like PredictIt and Polymarket.
-This comparison will help adjust for potential biases in polling, as betting markets can reveal hidden preferences or voter reluctance to openly state their choices.
-•	Modeling: We will use Bayesian hierarchical models to predict the election outcome at both the national and state levels.
-The model will integrate polling data, betting odds, and historical election trends.
-By comparing how betting and polling data align or diverge, we will refine the forecast, accounting for last-minute voter shifts or surprise outcomes.
-
-6. Poll-Betting Market Discrepancy
-
-A major component of this methodology is addressing the discrepancies between polling and betting markets.
-Our hypothesis is that participants are more likely to provide accurate predictions when they have a financial stake in the outcome.
-
-•	Betting behavior analysis: We will analyze how betting behavior correlates with stated voter preferences, identifying whether betting markets capture sentiment that traditional polling does not.
-•	Incentive impact: By comparing the results of those who opted to bet versus those who did not, we can determine whether financial incentives lead to more accurate election predictions.
-
-7. Budget Breakdown
-
-The $100,000 budget will be allocated as follows:
-
-•	Respondent recruitment: $10,000
-•	Survey administration (online and telephone): $20,000
-•	Betting pool incentives: $50,000 (for those who accurately predict the election outcome)
-•	Data validation and analysis: $10,000
-•	Poll aggregation and modeling: $10,000
-
-Conclusion
-
-This methodology uses a novel approach to incentivize accurate predictions by allowing respondents to bet on the election outcome.
-By combining traditional polling techniques with betting incentives, we aim to close the gap between polling data and betting market predictions, potentially uncovering hidden voter behaviors or biases that standard polling methods may miss.
-
-The survey can be found here:
-https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0
+This approach, inspired by [@Stantcheva’s guide on survey creation](https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0), leverages financial incentives to align respondent predictions with their genuine expectations.
+By combining innovative sampling, incentivized engagement, and rigorous data validation, this methodology aims to bridge the gap between traditional polling and betting market predictions, ultimately enhancing the accuracy of our election forecast.
 

### 2024-10-30 paper/paper.qmd
reformat the paper so that it looks better

@@
 The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
-The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages...
+The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...
 
@@
 
-# 2.1 The Dataset
+# 2.2 The Dataset
 
@@
 By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
-Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
+Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @FiveThirtyEight].
 In figure 3 we see the ammount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
@@
 
- # 2.2 Variables of Interest
+# 2.3 Variables of Interest
 In our logistic regression model predicting the 2024 U.S.
@@
 
-Model Overview
+## Model Overview
 
@@
 
-•	 $P(Y=1 | X)$  represents the probability of Trump winning (i.e.,  Y=1 ).
-•	 $\beta_0$  is the intercept term.
-•	 $\beta_1$  and  $\beta_2$  are the coefficients for the predictor variables, which include:
-•	 X_1 : The average polling percentage for Trump ( \text{trump\_pct} ).
-•	 X_2 : The average polling percentage for Harris ( \text{harris\_pct} ).
+•	 $P(Y=1 | X)$  represents the probability of Trump winning (i.e.,  Y=1 ). \newline
+•	 $\beta_0$  is the intercept term. \newline
+\newline $\beta_1$  and  $\beta_2$  are the coefficients for the predictor variables, which include: \newline
+•	 X_1 : The average polling percentage for Trump ( \text{trump\_pct} ). \newline
+•	 X_2 : The average polling percentage for Harris ( \text{harris\_pct} ). \newline
 
-Data and Features
+## Data and Features
 
@@
 
-	•	trump_pct: The percentage of respondents supporting Trump.
-	•	harris_pct: The percentage of respondents supporting Harris.
-	•	sample_size: The size of each poll, which influences the weight of that poll in our analysis.
+•	trump_pct: The percentage of respondents supporting Trump. \newline
+•	harris_pct: The percentage of respondents supporting Harris. \newline
+•	sample_size: The size of each poll, which influences the weight of that poll in our analysis. \newline
 •	start_date: The date when the poll began, which is used to calculate recency.
 
-Weight Calculation
+## Weight Calculation
 
@@
 
-Recency Weight:
+### Recency Weight:
 
@@
 
-Total Weight:
+### Total Weight:
 
@@
 
-Logistic Regression Implementation
+### Logistic Regression Implementation
 
-The logistic regression model is implemented using the LogisticRegression class from the sklearn library. The predictor variables are defined as:
+The logistic regression model is implemented using the LogisticRegression class from the sklearn library. The predictor variables are defined as: \newline
 
@@
 
-	•	Linearity: The log-odds of the outcome variable are linearly related to the predictor variables.
+•	Linearity: The log-odds of the outcome variable are linearly related to the predictor variables. \newline
 •	Independence: Observations are independent of one another.
 
-Limitations
+### Limitations
 
@@
 Polls with lower numeric grades or transparency scores are excluded, potentially introducing bias if high-quality polls are not representative of the entire electorate.
+\newline
 2.	Dynamic Voter Sentiment: Voter preferences can change rapidly, particularly in the lead-up to the election, and the model may not fully capture these shifts if not updated frequently.
 
-Alternative Models Considered
+### Alternative Models Considered
 
@@
 While Bayesian models allow for the incorporation of prior distributions and might provide additional insights through credible intervals, they require more complex implementation and careful selection of priors.
-Given the nature of the data and the objective of clear interpretability, the logistic regression model was chosen for its straightforward application and interpretability.
+Given the nature of the data and the objective of clear interpretability, the logistic regression model was chosen for its straightforward application and interpretability.\newline
 
-Conclusion
-
-The logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data.
+Ultimately the logistic regression model provides a structured approach to predict the likelihood of Trump winning in various states, based on recent and high-quality polling data.
 By focusing on key features such as polling percentages and sample sizes, the model captures the competitive dynamics between candidates and offers valuable insights into potential election outcomes.
@@
 
-This approach, inspired by [@Stantcheva’s guide on survey creation](https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0), leverages financial incentives to align respondent predictions with their genuine expectations.
+This approach, inspired by Stantcheva’s guide on survey creation [@Stantcheva], leverages financial incentives to align respondent predictions with their genuine expectations.
 By combining innovative sampling, incentivized engagement, and rigorous data validation, this methodology aims to bridge the gap between traditional polling and betting market predictions, ultimately enhancing the accuracy of our election forecast.

### 2024-10-31 paper/paper.qmd
add results and edi some of the quato code

--- 2024-10-25 paper/paper.qmd
@@
 ---
+
+```{python}
+#| echo: false
+import pandas as pd
+import matplotlib.pyplot as plt
+import jupyter
+from tabulate import tabulate
+import numpy as np
+```
 
@@
 
-The structure of the paper is as follows
+The structure of the paper is as follows: Section 2 describes the data collection process, detailing the sources, variables of interest, and criteria for selecting high-quality polls.
+Section 3 introduces the logistic regression model used for prediction, explaining the variables, weights, and methodology applied to forecast the election outcome.
+In Section 4, we present and analyze the model’s predictions, highlighting the probabilities of candidate success across key states.
+Section 5 discusses the model’s implications, limitations, and the observed contrasts between polling data.
 
@@
 # 2.1 Measurement
+
 The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
-This dataset reflects potential outcomes for the 2024 U.S.
+This dataset provides predictions for candidate support for the 2024 U.S.
 Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific poll conducted by different polling organizations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
-
-It’s important to note that each pollster employs unique sampling techniques and methodologies, which influence their numeric grades and transparency scores.
-To ensure the integrity of our analysis, we selected only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
-This approach is used so that the data reflects a robust and credible representation of voter preferences.
+Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+the surveys utilized one or multiple of the following method for rsponse gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
+The diversity of the methods of collection supplies for a colorful array of polls capturing all facets of voters.
 
 The polling data is collected through carefully structured survey questions, designed to elicit clear responses regarding voter preferences.
-By aggregating these individual responses, we are able to derive state-level averages for both candidates.
-This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
+This transformation of general voter sentiment into specific data points enables us to analyse the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
 
@@
 
+A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
+States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
+Thus the election result is normally decided by which party wins just a handful of states in the electoral college system.
+By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
+Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @538].
+In figure 3 we see the ammount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
+
+```{python}
+# Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot
+states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois", "Maine", 
+          "Maryland", "Massachusetts", "Minnesota", "New Jersey", "New York", "Oregon", 
+          "Rhode Island", "Vermont", "Washington", "District of Columbia", "New Mexico", 
+          "New Hampshire", "Colorado", "Virginia", "Arizona", "Georgia", "Michigan", 
+          "Nevada", "North Carolina", "Pennsylvania", "Wisconsin", "Alabama", "Alaska", 
+          "Arkansas", "Florida", "Idaho", "Indiana", "Iowa", "Kansas", "Kentucky", 
+          "Louisiana", "Mississippi", "Missouri", "Montana", "Nebraska", "North Dakota", 
+          "Ohio", "Oklahoma", "South Carolina", "South Dakota", "Tennessee", "Texas", 
+          "Utah", "West Virginia", "Wyoming"]
+votes = [54, 7, 3, 4, 19, 4, 10, 11, 10, 14, 28, 8, 4, 3, 12, 3, 5, 4, 10, 13, 
+         11, 16, 15, 6, 16, 19, 10, 9, 3, 6, 30, 4, 11, 6, 6, 8, 8, 6, 10, 4, 5, 
+         3, 17, 7, 9, 3, 11, 40, 6, 4, 3]
+
+# Calculate cumulative votes
+cumulative_votes = np.cumsum(votes)
+cumulative_votes_reversed = np.cumsum(votes[::-1])
+x = np.arange(len(states))
+
+plt.figure(figsize=(18, 8))
+plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
+x_reversed = np.flip(x)
+plt.step(x_reversed, cumulative_votes_reversed, where='post', color='red', linewidth=2, marker='o')
+
+# Threshold for filling
+dem_threshold = np.argmax(cumulative_votes >= 227)
+plt.fill_between(x[:dem_threshold + 1], cumulative_votes[:dem_threshold + 1], step='post', color='blue', alpha=0.4)
+
+# Adding threshold line and markers
+plt.axhline(270, color='purple', linestyle='--', linewidth=2)
+plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
+plt.xlabel('Electoral College by State', fontsize=12)
+plt.ylabel('Total Votes', fontsize=12)
+plt.title('Cumulative Electoral Votes by State', fontsize=15)
+plt.grid(True, linestyle='--', alpha=0.5)
+plt.tight_layout()
+plt.show()
+
+```
+
+It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
+Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
+a good score reflecting low propensity to be impacted by errors and biases that can occur in survey sampling.
+It also accounts for impacts that might bias a survey's score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
+Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
+By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
+We use this single measurement out of 3 to ensure the integrity of our analysis.
+Selecting only those polls that originate from reputable pollsters, specifically targeting those with a numeric grade of 2.5 or higher.
+This approach is used so that the data reflects a robust and credible representation of voter preferences.
+
+In keeping with this aim, we focused our analysis on polls conducted after September.
+Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, for example Biden dropping out as a candidate on July 21st.
+These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
+By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
+
+```{python}
+#| echo: false
+
+# Load the data (replace with your CSV file path)
+df = pd.read_csv('../data/02-analysis_data/merged_swing_state_data.csv')
+
+# Group by state and calculate summary statistics
+summary_stats = df.groupby('state').agg(
+    numeric_grade_mean=('numeric_grade', 'mean'),
+    numeric_grade_std=('numeric_grade', 'std'),
+    numeric_grade_min=('numeric_grade', 'min'),
+    numeric_grade_max=('numeric_grade', 'max'),
+    numeric_grade_count=('numeric_grade', 'count'),
+    sample_size_mean=('sample_size', 'mean'),
+    sample_size_std=('sample_size', 'std'),
+    sample_size_min=('sample_size', 'min'),
+    sample_size_max=('sample_size', 'max'),
+    sample_size_count=('sample_size', 'count')
+).reset_index()
+
+# Reshape the summary table to long format
+summary_long = pd.melt(
+    summary_stats,
+    id_vars='state',
+    value_vars=[
+        'numeric_grade_mean', 'numeric_grade_std', 'numeric_grade_min', 'numeric_grade_max', 'numeric_grade_count',
+        'sample_size_mean', 'sample_size_std', 'sample_size_min', 'sample_size_max', 'sample_size_count'
+    ],
+    var_name='variable_stat', value_name='value'
+)
+
+# Split 'variable_stat' into 'variable' and 'stat'
+summary_long[['variable', 'stat']] = summary_long['variable_stat'].str.rsplit('_', n=1, expand=True)
+
+# Drop the original column
+summary_long.drop('variable_stat', axis=1, inplace=True)
+
+# Pivot the table to get desired output
+summary_pivot = summary_long.pivot_table(
+    index=['state', 'variable'], columns='stat', values='value', aggfunc='first'
+).reset_index()
+
+# Sort for better readability
+summary_pivot = summary_pivot.sort_values(by=['state', 'variable'])
+
+summary_pivot[['mean', 'std', 'min', 'max']] = summary_pivot[['mean', 'std', 'min', 'max']].round(1)
+
+# Create a list of rows for tabulate (merging state rows visually)
+rows = []
+last_state = None
+for _, row in summary_pivot.iterrows():
+    state = row['state']
+    variable = row['variable']
+    mean, std, min_, max_, count = row['mean'], row['std'], row['min'], row['max'], row['count']
+
+    # Only show state name on the first row of each group
+    if state == last_state:
+        state = ''
+    else:
+        last_state = state
+
+    rows.append([state, variable, mean, std, min_, max_, count])
+
+# Print the table using tabulate
+print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='fancy_grid'))
+```
+
  # 2.2 Variables of Interest
+In our logistic regression model predicting the 2024 U.S.
+Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data.
+The percentage of respondant favouring a candidate become the key variables for prediction.
+Within the collection of polls in our sample the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump , and `harris_pct` respectively.
+Polls ar taking towards the build up of an election and thus ar taken by repondants at different times.
+Graph  x shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
+Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
+```{python}
+#| echo: false
+
+# Load data (use an accurate path or dataset)
+data = pd.read_parquet('../data/02-analysis_data/merged_swing_state_data.parquet')
+data['end_date'] = pd.to_datetime(data['end_date'])
+
+plt.figure(figsize=(12, 6))
+plt.hist(data['end_date'], bins=20, edgecolor='black', alpha=0.7)
+plt.xlabel('Poll End Date')
+plt.ylabel('Poll Count')
+plt.title('Poll Distribution Over Time')
+plt.xticks(rotation=45)
+plt.tight_layout()
+plt.show()
+
+```
+
+Another aspects of the polls we wish to consider is the population size of each.
+Larger polls result in a stronger reduction in randomness in the response.
+The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers.
+In table x we can see the variety within the states polls.
+For use in the model the percentages in these polls is weighted for recenecy and size of the population.
 
@@
  # 4 Results {#sec-results}
-The results of our analysis are as follows.
-We have Donald Trump predicted to win Arizona, Georgia and Michigan for a total of 42 of the 93 swing state votes.
-While Kamala Harris is projected to win Nevada, North Carolina, Pennsylvania and Wisconsin for the remaining 51 swing state votes.
-This results in a win for the democrats and the first female US president!
+ ## Regression results
+
+ ```{Python}
+#| echo: false
+
+import pandas as pd
+from tabulate import tabulate
+
+# Data setup
+data = {
+    'State': ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 'Pennsylvania', 'Wisconsin'],
+    'Trump % (Mean)': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
+    'Trump % (SD)': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
+    'Harris % (Mean)': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
+    'Harris % (SD)': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
+    'Percentage Chance of Trump Winning': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
+}
+
+# Create DataFrame
+df = pd.DataFrame(data)
+
+# Display as a professional table
+table_output = tabulate(
+    df, headers='keys', tablefmt='grid', showindex=False,
+    floatfmt=(".2f", ".2f", ".2f", ".2f", ".2f", ".2f")
+)
+print(table_output)
+
+```
+
+in figure xx we cans see th results from the model described.
+
+In Arizona and Georgia, the model predicts the strongest likelihood of a Trump victory, with a mean polling percentage showing Trump over 2 percentage points ahead in Arizona.
+In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.The data is thus resulting in an strong 86.85% and and 71.18% likelyhood in Trump winning in Arizona and Georgia respectively.
+Trump is also predicted to win in Michigan, alhough by smaller margins as shown in figure xx.
+As a rssult the model predicts trump winning with only a 53% chance.
+With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.
+
+```{python}
+#| echo: false
+
+# Data setup (can adjust based on model results)
+data = {
+    'State': ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 'Pennsylvania', 'Wisconsin'],
+    'Trump % (Mean)': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
+    'Trump % (SD)': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
+    'Harris % (Mean)': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
+    'Harris % (SD)': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
+    'Trump Win Probability (%)': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
+}
+df = pd.DataFrame(data)
+df['Distance from 50%'] = df['Trump Win Probability (%)'] - 50
+
+# Centered bar chart
+fig, ax = plt.subplots(figsize=(10, 6))
+colors = ['red' if x > 0 else 'blue' for x in df['Distance from 50%']]
+ax.barh(df['State'], df['Distance from 50%'], color=colors)
+ax.axvline(0, color='black')
+ax.set_xlabel('Distance from 50% (Trump Winning Probability)')
+ax.set_title('Probability of Trump Win, Centered from 50%')
+plt.tight_layout()
+plt.show()
+
+```
+
+Kamala is predicted to win he remaining of th states and is givn fairly high chances of locking these up.
+despite only winning Nevada, North Carolina and Pennsilvania by under a percntag point in the model Kamalas probablity of winning thos eelectoral votes is vry high.
+This is due to the small standard dviations shown in table x making he prdicted rul of the states leaning democrat being mor likely.
+Wisconson is shown to be her srongest and most likley victory of the swing states. 
+Overall this givs Harris 51 of the 93 swing state votes his results in a win for the democrats and the first female US president!
+
+## Trump overall probability
+-- Taking he probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and loses for Trump in these swing states adn deduce his overall probability of reaching 270 electoral votes.
+Basd on the modles probabilities his current expected votes is 41.05, below the threshold of 51 he needs based on our assumptions.
+
+Seperating the outcomes into those that give trump the required electoral college votes, we can obtain  probability that trump ceizes a second term.
+
+Using a probaboiility theory, and assuming independence among the states, in our approach.
+When accounting for all possible state outcome combinations, the model calculates that Trump has approximately a 22.5% chance of securing at least 51 electoral votes from these key swing states.
+This probability aggregates the likelihoods across all combinations where the sum of electoral votes meets or exceeds the 51-vote threshold.
+The analysis reveals that despite high individual probabilities in certain states, the overall chance of Trump amassing sufficient electoral votes from the swing states is less than one in four.
 
@@
 
-The survey can be found here:
+An example of the survey can be found here:
 https://docs.google.com/forms/d/e/1FAIpQLSc5LYCAd0OmiIc-LHRFUVgVGuSMKzayVWZB7VOrsrDOe5742w/viewform?vc=0&c=0&w=1&flr=0
+
+# appndix B 
+say what kind of ampling. 
+reference rsearch on wha the sampling mthod is
+
+## what is included in the poll?
+
+## Methodology and why?
+
 # References

### 2024-10-31 paper/paper.qmd
prepare for merge

--- 2024-10-30 paper/paper.qmd
@@
 from tabulate import tabulate
+import numpy as np
 ```
@@
 Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
+the surveys utilized one or multiple of the following method for rsponse gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
+The diversity of the methods of collection supplies for a colorful array of polls capturing all facets of voters.
 
@@
 ```{python}
-import matplotlib.pyplot as plt
-import numpy as np
+# Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot
+states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois", "Maine", 
+          "Maryland", "Massachusetts", "Minnesota", "New Jersey", "New York", "Oregon", 
+          "Rhode Island", "Vermont", "Washington", "District of Columbia", "New Mexico", 
+          "New Hampshire", "Colorado", "Virginia", "Arizona", "Georgia", "Michigan", 
+          "Nevada", "North Carolina", "Pennsylvania", "Wisconsin", "Alabama", "Alaska", 
+          "Arkansas", "Florida", "Idaho", "Indiana", "Iowa", "Kansas", "Kentucky", 
+          "Louisiana", "Mississippi", "Missouri", "Montana", "Nebraska", "North Dakota", 
+          "Ohio", "Oklahoma", "South Carolina", "South Dakota", "Tennessee", "Texas", 
+          "Utah", "West Virginia", "Wyoming"]
+votes = [54, 7, 3, 4, 19, 4, 10, 11, 10, 14, 28, 8, 4, 3, 12, 3, 5, 4, 10, 13, 
+         11, 16, 15, 6, 16, 19, 10, 9, 3, 6, 30, 4, 11, 6, 6, 8, 8, 6, 10, 4, 5, 
+         3, 17, 7, 9, 3, 11, 40, 6, 4, 3]
 
-# Step 1: Define the Electoral Votes for Each State
-states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois",
-"Maine", "Maryland", "Massachusetts", "Minnesota", "New Jersey",
-"New York", "Oregon", "Rhode Island", "Vermon", "Washington",
-"District of Columbia", "New Mexico", "New Hampshire", "Colorado", "Virginia",
-"Arizona",
-"Georgia",
-"Michigan",
-"Nevada",
-"North Carolina",
-"Pennsylvania",
-"Wisconsin",
-"Alabama", "Alaska", "Arkansas", "Florida", "Idaho",
-"Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana",
-"Mississippi", "Missouri", "Montana", "Nebraska", "North Dakota",
-"Ohio", "Oklahoma", "South Carolina", "South Dakota", "Tennessee",
-"Texas", "Utah", "West Virginia", "Wyoming"]
-print(len(states))
-votes = [54, 7 , 2, 4, 19, 4, 10, 11, 10, 14, 28, 8, 4, 3, 12, 3, 5, 4, 10, 13,
-         11,16,15,6,16,19,10,
-         9,3,6,30,4,11,6,6,8,8,6,10,4,5,3,17,7,9,3,11,40,6,4,3]
+# Calculate cumulative votes
+cumulative_votes = np.cumsum(votes)
+cumulative_votes_reversed = np.cumsum(votes[::-1])
+x = np.arange(len(states))
 
-print(len(votes))
-# Step 2: Compute the Cumulative Electoral Votes
-cumulative_votes = np.cumsum(votes)
-
-# Compute Cumulative Votes in Reverse Order for Republicans
-cumulative_votes_reversed = np.cumsum(votes[::-1])
-
-# X-axis values (state index)
-x = np.arange(len(cumulative_votes))
-
-# Step 3: Plot the Step Function for Democrats (left to right)
 plt.figure(figsize=(18, 8))
 plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
-
-# Reverse the x-axis for the Republican step function (right to left)
 x_reversed = np.flip(x)
@@
 
-# Step 4: Customize the Plot
-plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
-plt.yticks(np.arange(0, 550, 50))
+# Threshold for filling
+dem_threshold = np.argmax(cumulative_votes >= 227)
+plt.fill_between(x[:dem_threshold + 1], cumulative_votes[:dem_threshold + 1], step='post', color='blue', alpha=0.4)
 
-# Fill the area under the curve with blue for Democrats (left to right)
-threshold_index = np.argmax(cumulative_votes >= 227)  # First point where cumulative votes reach 270
-plt.fill_between(x[:threshold_index + 1], cumulative_votes[:threshold_index + 1], 
-                 step='post', color='blue', alpha=0.4)
-
-# Fill the area under the curve with red for Republicans (right to left)
-threshold_index = np.argmax(cumulative_votes >= 219)
-plt.fill_between(x_reversed[:threshold_index + 5], cumulative_votes_reversed[:threshold_index + 5], step='post', color='red', alpha=0.4)
-
-# Add a horizontal line at 270 votes (threshold to win)
+# Adding threshold line and markers
 plt.axhline(270, color='purple', linestyle='--', linewidth=2)
 plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
-
-# Add lines for current Democratic and Republican thresholds
-plt.axhline(225, color='blue', linestyle='--', linewidth=1.5)
-plt.text(210, 220, '225 votes', horizontalalignment='center', color='blue', fontsize=12)
-
-# Add Labels and Title
-plt.xlabel('Electoral College Markets (electoral votes)', fontsize=12)
-plt.ylabel('Total Electoral Votes', fontsize=12)
-plt.title('Step Function of Total Electoral Votes\nwith States Ordered by Vote Share', fontsize=15)
-
-# Add Grid for Better Visibility
+plt.xlabel('Electoral College by State', fontsize=12)
+plt.ylabel('Total Votes', fontsize=12)
+plt.title('Cumulative Electoral Votes by State', fontsize=15)
 plt.grid(True, linestyle='--', alpha=0.5)
-
-# Adjust Layout for Better Fit
 plt.tight_layout()
-
-# Step 5: Show the Plot
 plt.show()
@@
 
-# Load your data (replace with your actual data file path)
+# Load data (use an accurate path or dataset)
 data = pd.read_parquet('../data/02-analysis_data/merged_swing_state_data.parquet')
-
-# Ensure date column is in datetime format
 data['end_date'] = pd.to_datetime(data['end_date'])
 
-# Plot the histogram of poll quantity by end date
 plt.figure(figsize=(12, 6))
 plt.hist(data['end_date'], bins=20, edgecolor='black', alpha=0.7)
-
-# Adding labels and title
-plt.xlabel('End Date of Poll')
-plt.ylabel('Number of Polls')
-plt.title('Histogram of Poll Quantity by End Date')
+plt.xlabel('Poll End Date')
+plt.ylabel('Poll Count')
+plt.title('Poll Distribution Over Time')
 plt.xticks(rotation=45)
-
-# Show the plot
 plt.tight_layout()
@@
 
-# Appendix A: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting
+# Apendix
+(include some detail on how we are getting rid of biass. and focus on how we are spcfically uing stratifid and clustr sampling. how we will reach ou. how we will gt rsponce.)
+
+## 1. Sampling Approach
+
+To ensure that our sample represents the diversity of the US electorate and minimizes potential biases, we employ stratified and cluster sampling techniques.
+With a budget of $100,000, our goal is to gather data from 5,000 respondents, focusing on key swing states while preserving national representativeness.
+The sample will be stratified by:
 
@@
 
+## what is included in the poll?
+
+## Methodology and why?
+
 # References

### 2024-11-01 paper/paper.qmd
Appndix B first Draft

@@
 
+The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
+The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...
 The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
@@
 
-# Apendix
-(include some detail on how we are getting rid of biass. and focus on how we are spcfically uing stratifid and clustr sampling. how we will reach ou. how we will gt rsponce.)
-
-## 1. Sampling Approach
-
-To ensure that our sample represents the diversity of the US electorate and minimizes potential biases, we employ stratified and cluster sampling techniques.
-With a budget of $100,000, our goal is to gather data from 5,000 respondents, focusing on key swing states while preserving national representativeness.
-The sample will be stratified by:
+# Appendix A: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting
 
@@
 
-# appndix B 
-say what kind of ampling. 
-reference rsearch on wha the sampling mthod is
+# Appendix B: Siena / New York Tims Pennsilvania poll
+
+This section will discuss a particular poll from the database of polls used in the above analysis, one made by sienna and the new york times.
+The New York Times partners with the Siena college in order to carry out political polls throughout the election period, these are often high quality polls according to their rating, likely because the nyt has been doing these polls for a long time.
+The NYT polls often emphasise a focus on the battleground states and poll then more than others making these very useful in this paper’s analysis.
+This action will focus on one poll done in Philadelphia, but those in other states follow the same methodology and patterns.
+This consistency is important for the NYT when they are reporting their results, as it makes comparison around the US viable and reliable.
+In going over the methodology this appendix aims to understand what about this pollster makes them high quality and what might they leave wanting.
+## Sampling method
+The poll employs a response rate-adjusted stratified sampling method using the L2 voter file, which contains detailed demographic information on registered voters.
+Initially the interviewee's are selected randomly from a national list of registered voters, response are then weighted.
+The stratification, often referred to as weighing, is applied to the random sample collected in order to properly reflect the entire demographic.
+The stratification accounts for:
+- Statehouse district
+- Political party affiliation
+- Race
+- Gender
+- Marital status
+- Household size
+- Turnout history
+- Age
+- Home ownership
+The sampling is conducted in order to address telephone coverage around the state as well as nonresponse rates.
+Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Phyladelphia,.
+The voter file is then ratified and weights are based on historically modelled response ra and coverage.
+
+## Collection
+Times/ Siena polls are conducted by telephone, using live interviews.
+About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish.
+In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (https://www.nytimes.com/interactive/2024/09/19/us/politics/times-siena-inquirer-poll-pennsylvania-likely-electorate.html).
+This high call volume is key to maximising response from all groups.
+In a further effort to get a good sample additional effort is made to constant under-represented groups, for example people with lower education levels.
+
+## Questionnaire design
+The questions themselves included more than just the information on who a participant would vote on.
+Followup questions were given either because or regardless of what their previous answers were.
+One example would be follow up questions for those who answer as “don’t know/ refused” in order to understand some reasoning behind this answer.
+Those who did select an option also were asked questions to get a clear image of their sentiment or reasoning.
+The benefit of these additional questions is to bring context to the initial binary question.
+
+## Methodology pros and cons
+The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on it's own more redundant.
+(https://hdsr.mitpress.mit.edu/pub/ejk5yhgv#n0xinrk8s1g).
+This is clearly a factor here as the response rate was below 2%.
+The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues [@Brick2013Nonresponse].
+The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population.
+Making the means generate more accurate.
+With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias.
+It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias.
+Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 [Clinton t al (2021) kennedy et al (2018)].
+With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.
+The benefit of how the survey was taken is the detail added from the additional questions asked, some key questions allow for further stratification based on the answers.
+For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are.
+One documented phenomenon worth mentioning is social desirability bias, that people do tend to choose to answer {definition] (cite).
+This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote ().
+The pollster’s use of both modelled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically.
+The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however siena/NYT took a methodology that aims to negate this as much as possible.
+Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides.
+Qualitative insights to voters' choices or indecision creates a more robust and significant dataset.
+By comparison this is far more valuable than a singular binary choice quiz you may see online.
+
+In conclusion, while the pollster's methodology is robust and incorporates several best practices in survey research, inherent challenges such as low response rates and the possibility of nonresponse and social desirability biases must be acknowledged.
+This poll’s methodology makes great efforts to mitigate these biases through its weighting methods, yet these factors could affect the accuracy of the poll results.
+Nonetheless, The poll provides valuable insights and remains a useful resource for analysing voter preferences, especially due to its aforementioned weighting strategy.
+The poll's transparency in methodology makes it an even more reliable resource and enhances the credibility and comparability of the findings.
 

### 2024-11-02 paper/paper.qmd
Paper edits

@@
 ```{python}
+#| ccho: false
+#| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, accordng to 538.
+The chart distinguishes between Democrat, Republican, and swing states, with a color-coded fill under each party’s cumulative curve.
+Key threshold of 270 electoral vote mark needed to secure a victory is diplayed.
+This visualization shows the democrats lead with 227 votes and the Republicans with 219 vot secured"
 # Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot
+
 states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois", "Maine", 
@@
 plt.fill_between(x[:dem_threshold + 1], cumulative_votes[:dem_threshold + 1], step='post', color='blue', alpha=0.4)
+threshold_index = np.argmax(cumulative_votes >= 219)
+plt.fill_between(x_reversed[:threshold_index + 5], cumulative_votes_reversed[:threshold_index + 5], step='post', color='red', alpha=0.4)
 
@@
 plt.ylabel('Total Votes', fontsize=12)
-plt.title('Cumulative Electoral Votes by State', fontsize=15)
+plt.title('Cumulative Electoral Votes by State', fontsize=25)
 plt.grid(True, linestyle='--', alpha=0.5)
@@
 ```{python}
+
 #| echo: false
-
+#| fig-cap: "Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states."
 # Load the data (replace with your CSV file path)
@@
 
-# Split 'variable_stat' into 'variable' and 'stat'
+# Rename elements
 summary_long[['variable', 'stat']] = summary_long['variable_stat'].str.rsplit('_', n=1, expand=True)
+summary_long['variable'] = summary_long['variable'].str.replace('_', ' ', regex=False).str.title()
 
@@
 # Print the table using tabulate
-print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='fancy_grid'))
+print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='simple_grid'))
 ```
+**Table 1: Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states.**
 
@@
 #| echo: false
+#| fig-cap: "A histogram of the poll quantities, aggregated by the end date of each poll, shows the distribution of polling data over time.
+This chart reveals the frequency of polls leading up to the election and provides insight into polling activity trends, with potential spikes reflecting periods of increased public interest or campaign activity."
 
@@
 plt.ylabel('Poll Count')
-plt.title('Poll Distribution Over Time')
+plt.title('Poll Distribution Over Time', fontsize=25)
 plt.xticks(rotation=45)
@@
 
- ```{Python}
+```{python}
 #| echo: false
-
-import pandas as pd
-from tabulate import tabulate
-
 # Data setup
@@
     'State': ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 'Pennsylvania', 'Wisconsin'],
-    'Trump % (Mean)': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
-    'Trump % (SD)': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
-    'Harris % (Mean)': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
-    'Harris % (SD)': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
-    'Percentage Chance of Trump Winning': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
+    'Trump %': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
+    'Trump SD': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
+    'Harris %': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
+    'Harris SD': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
+    'Trump Winning %': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
 }
@@
 # Display as a professional table
-table_output = tabulate(
+print(tabulate(
     df, headers='keys', tablefmt='grid', showindex=False,
     floatfmt=(".2f", ".2f", ".2f", ".2f", ".2f", ".2f")
-)
-print(table_output)
+))
 
 ```
+**Table 2: Results table for the model describd above.
+Show support fo trump and Kamala at the stat level accordng to avalable polls.
+Small standard dviation show strong results, yet the rac is very close in many states.
+Percentages above 50% show the favoured candidate in each state.**
 
-in figure xx we cans see th results from the model described.
+in figure xx we cans see the results from the model described.
 
@@
 #| echo: false
-
+#| fig-cap: "A centered bar chart where each state has a bar reflecting Trump’s probability of winning.
+The bars extend to the right or left from the center line at 50%, with the distance representing the magnitude of deviation from an even chance.
+This chart visualizes Trump’s relative strength in each swing state and highlights that he s only favoured in Georgia and Arizona."
 # Data setup (can adjust based on model results)
@@
 ax.set_xlabel('Distance from 50% (Trump Winning Probability)')
-ax.set_title('Probability of Trump Win, Centered from 50%')
+ax.set_title('Probability of Trump Win in swing states, Centered from 50%', fontsize=25)
 plt.tight_layout()
@@
 First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals).
-Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve non-linear influences that are difficult to predict.
+As polling results are based on public opinion, large shock events are likely and expected, moreover scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modeled here.
+Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve nonlinear influences that are difficult to predict.
 
@@
 Training the model on a more comprehensive dataset that includes these states and other predictive variables, such as demographic factors or economic indicators, could lead to a more accurate forecast.
+
+Candidate Trump’s low likelihood of winning shown by our results may under-represent his support, this claim is supported by the disparity with the betting markets.
+This could be a result of a common problem with polling data, response rate bias.
+In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day.
+Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support Clinton t al (2021) kennedy et al (2018)].
+Should this trend continue, and be present in the polls used in this analysis, it may be expected that Trump could surprise the model and win some of these swing states.
+Another reason Trump could outperform the model is another bias often seen in surveys, one of social desirability.
+Trump is a unique candidate as, if elected, he would become the first president to be elected with federal crime conviction.
+This along with other negative publicity the former host of tv’s apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support.
+Even with the quality polls used in this analysis, where steps are taken to ensure representative results(see appendix B), we could be observing a skew in the results.
 

### 2024-11-02 paper/paper.qmd
File Cleanup and Parquet impelmentation

@@
 #| fig-cap: "Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states."
-# Load the data (replace with your CSV file path)
-df = pd.read_csv('../data/02-analysis_data/merged_swing_state_data.csv')
+
+df = pd.read_parquet('../data/02-analysis_data/merged_swing_state_data.parquet')
 

### 2024-11-02 scripts/00-simulate_data.py

@@ -70,5 +70,5 @@ df = pd.DataFrame(data, columns=[
 # Display the first few rows of the simulated dataset
 print(df.head())
 
-# Save the simulated dataset to a CSV file
-df.to_csv("data/00-simulated_data/simulated_data.csv", index=False)
+# Save the simulated dataset to a parquet file
+df.to_parquet("data/00-simulated_data/simulated_data.parquet", index=False)

### 2024-11-02 scripts/01-test_simulated_data.py

@@ -12,7 +12,7 @@ import pandas as pd
 import unittest
 
 # Load the simulated dataset
-df = pd.read_csv("simulated_polling_data.csv")
+df = pd.read_parquet("simulated_polling_data.parquet")
 
 class TestSimulatedPollingData(unittest.TestCase):
 

### 2024-11-02 scripts/03-clean_data.py

@@ -63,7 +63,7 @@ print(df_swing['unique_id'].nunique())
 
 print(df_swing["numeric_grade"].min())
 # save raw swing stat data
-df_swing.to_parquet("data/02-analysis_data/swing_state_polls.csv")
+df_swing.to_parquet("data/02-analysis_data/swing_state_polls.parquet")
 
 #creating data to be analysed (grouping by )
 df_trump = df_swing[df_swing['candidate_name'] == "Donald Trump"]

### 2024-11-02 scripts/04-test_analysis_data.py

@@ -12,7 +12,7 @@
 import pandas as pd
 
 # Load the dataset
-df = pd.read_csv('data/02-analysis_data/merged_swing_state_data.parquet')
+df = pd.read_parquet('data/02-analysis_data/merged_swing_state_data.parquet')
 
 def test_trump_harris_pct_sum():
     """

### 2024-11-02 scripts/06-model.py

@@ -2,7 +2,7 @@ import pandas as pd
 from sklearn.linear_model import LogisticRegression
 
 # Load the cleaned polling data
-clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")
+clean_poll_data = pd.read_parquet("data/02-analysis_data/merged_swing_state_data.parquet")
 
 # Create a weight for each poll based on recency and sample size
 # For simplicity, let's assume that the weight is based on the sample size directly

### 2024-11-02 scripts/06-model_data.py

@@ -1,31 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates election polling data
-# Author: Alexander Guarasci & Jacob Gilbert
-# Date: November 4, 2024
-# Contact: alexander.guarasci@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-  # - `pandas` must be installed (pip install polars)
-  # - `numpy` must be installed (pip install numpy)
-  # - `sklearn` must be installed (pip install sklearn)
-
-#### Workspace setup ####
-import pandas as pd
-import numpy as np
-from sklearn.model_selection import train_test_split
-from sklearn.linear_model import LogisticRegression
-from sklearn.metrics import classification_report, accuracy_score, confusion_matrix
-
-# Load the filtered swing state data
-df_swing = pd.read_csv("data/raw_data/swing_state_polls.csv")
-
-# Step 1: Create Binary Target (Trump Win or Lose)
-# Pivot the data to compare Trump's and Harris's polling percentages side by side
-swing_pivot = df_swing.pivot_table(
-    values='pct', 
-    index=['state', 'poll_end_date', 'poll_id'],
-    columns='candidate_name'
-).reset_index()
-
-print(df_swing['candidate_name'].unique())
-# Drop any rows with missing values for Trump or Harris

### 2024-11-02 scripts/graph_modelling/poll_date_hist.py

@@ -2,7 +2,7 @@ import pandas as pd
 import matplotlib.pyplot as plt
 
 # Load the dataset
-data = pd.read_csv("data/02-analysis_data/swing_state_polls.csv")  # Replace with your file path
+data = pd.read_parquet("data/02-analysis_data/swing_state_polls.parquet")  # Replace with your file path
 
 # Convert the end_date to datetime for plotting
 data['end_date'] = pd.to_datetime(data['end_date'])

### 2024-11-02 scripts/graph_modelling/summary_table.py

@@ -1,8 +1,8 @@
 import pandas as pd
 from tabulate import tabulate
 
-# Load the data (replace with your CSV file path)
-df = pd.read_csv('data/02-analysis_data/merged_swing_state_data.csv')
+# Load the data (replace with your parquet file path)
+df = pd.read_parquet('data/02-analysis_data/merged_swing_state_data.parquet')
 
 # Group by state and calculate summary statistics
 summary_stats = df.groupby('state').agg(

### 2024-11-02 scripts/graph_modelling/vote_pct_mean.py

@@ -1,8 +1,7 @@
 import pandas as pd
 
 # Assuming df is your DataFrame with the relevant data
-# Replace 'path_to_your_csv' with the actual path if reading from CSV
-# df = pd.read_csv('path_to_your_csv')
+
 
 df = pd.read_parquet('data/02-analysis_data/merged_swing_state_data.parquet')
 

### 2024-11-02 scripts/graphs.py

@@ -4,8 +4,8 @@ import matplotlib.pyplot as plt
 import seaborn as sns
 
 # Load the merged swing state polling data
-data_path = "data/02-analysis_data/merged_swing_state_data.csv"
-df = pd.read_csv(data_path)
+data_path = "data/02-analysis_data/merged_swing_state_data.parquet"
+df = pd.read_parquet(data_path)
 
 # Convert date columns to datetime format
 df['start_date'] = pd.to_datetime(df['start_date'])

### 2024-11-03 paper/paper.qmd
more paper edits

@@
 
-This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris.
-By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump or Harris victory in each state.
-This approach allows us to analyze voter support patterns across the states and predict election outcomes more accurately.
+This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likly decide th elections result.
+By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump's victory in each state.
+This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.
 
-The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, derived from aggregated state-level polling averages.
+The estimand in this study is the probability that Donald Trump wins a given state, derived from aggregated state-level polling averages.
 The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, with predictor variables comprising the weighted average polling percentages for both candidates.
@@
 Accurate election predictions provide valuable insights into voter dynamics, helping political analysts, campaigns, and the public anticipate electoral outcomes.
-By focusing on high-quality polling data and incorporating weights for recency and sample size, our model enhances prediction reliability and identifies key regions where voter sentiment may shift, ultimately impacting the election..
+By focusing on high-quality polling data and incorporating weights for recency and sample size, our model enhances prediction reliability and identifies key regions where voter sentiment may shift, ultimately impacting the election.
 
@@
 By aggregating these individual responses, we are able to derive state-level sentiment for both candidates.
-This transformation of general voter sentiment into specific data points enables us to analyse the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
+This transformation of general voter sentiment into specific data points enables us to analyze the competitive landscape between Trump and Harris across swing states, providing a clearer understanding of electoral dynamics as they relate to the upcoming election.
 
@@
 A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
-States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
+States that do not fall under this definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
 Thus the election result is normally decided by which party wins just a handful of states in the electoral college system.
 By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
-Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @FiveThirtyEight].
-In figure 3 we see the ammount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
+Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin; dtermined to be swing states in 2024 [@nyt_swing, @FiveThirtyEight].
+In figure 1 we see the current distribution of electoral votes that each party hold according to those sources as of October 20th, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
 
@@
 
-plt.figure(figsize=(18, 8))
+plt.figure(figsize=(18, 14))
 plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
@@
 # Adding threshold line and markers
+plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
 plt.axhline(270, color='purple', linestyle='--', linewidth=2)
@@
 Polls ar taking towards the build up of an election and thus ar taken by repondants at different times.
-Graph  x shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
+Figur 2 shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
 Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
@@
 The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers.
-In table x we can see the variety within the states polls.
+In table 1 we can see the variety within the states polls.
 For use in the model the percentages in these polls is weighted for recenecy and size of the population.
@@
 
-in figure xx we cans see the results from the model described.
-
 In Arizona and Georgia, the model predicts the strongest likelihood of a Trump victory, with a mean polling percentage showing Trump over 2 percentage points ahead in Arizona.
 In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.The data is thus resulting in an strong 86.85% and and 71.18% likelyhood in Trump winning in Arizona and Georgia respectively.
-Trump is also predicted to win in Michigan, alhough by smaller margins as shown in figure xx.
-As a rssult the model predicts trump winning with only a 53% chance.
+Trump is also predicted to win in Michigan, although by smaller margins as shown in this state.
+As a result the model predicts trump winning with only a 53% chance, showing his odds are near to a coin flip.
 With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.
@@
 
-Kamala is predicted to win he remaining of th states and is givn fairly high chances of locking these up.
-despite only winning Nevada, North Carolina and Pennsilvania by under a percntag point in the model Kamalas probablity of winning thos eelectoral votes is vry high.
-This is due to the small standard dviations shown in table x making he prdicted rul of the states leaning democrat being mor likely.
-Wisconson is shown to be her srongest and most likley victory of the swing states. 
-Overall this givs Harris 51 of the 93 swing state votes his results in a win for the democrats and the first female US president!
+Kamala is predicted to win the remaining states and is given fairly high chances of locking these up.
+despite only winning Nevada, North Carolina and Pennsilvania by under a percntage point in the model Kamala's probablity of winning those electoral votes is very high.
+This is due to the small standard deviations shown in table 2 making the predicted outcome of a victory more secure.
+Wisconson is shown to be her srongest and most likley victory of the swing states, with Trump only having a 15% chance here. 
+Overall this givs Harris 51 of the 93 swing state votes, our model expecting a win for the Democrats and the first female US president!
 
 ## Trump overall probability
--- Taking he probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and loses for Trump in these swing states adn deduce his overall probability of reaching 270 electoral votes.
-Basd on the modles probabilities his current expected votes is 41.05, below the threshold of 51 he needs based on our assumptions.
+Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and loses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes.
+Based on the modeled probabilities his current expected votes is 41.05, below the threshold of 51 he needs based on our assumptions.
 
-Seperating the outcomes into those that give trump the required electoral college votes, we can obtain  probability that trump ceizes a second term.
-
-Using a probaboiility theory, and assuming independence among the states, in our approach.
+Seperating the outcomes into those that give trump the required electoral college votes, we can obtain probability that trump seizes a second term.
+Using probability theory, and assuming independence among the states outomes, in our approach.
 When accounting for all possible state outcome combinations, the model calculates that Trump has approximately a 22.5% chance of securing at least 51 electoral votes from these key swing states.
@@
 In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day.
-Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support Clinton t al (2021) kennedy et al (2018)].
+Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support [@clinton2021task, @kennedy2018evaluation].
 Should this trend continue, and be present in the polls used in this analysis, it may be expected that Trump could surprise the model and win some of these swing states.
@@
 This along with other negative publicity the former host of tv’s apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support.
-Even with the quality polls used in this analysis, where steps are taken to ensure representative results(see appendix B), we could be observing a skew in the results.
+Even with the quality polls used in this analysis, where steps are taken to ensure representative results (see appendix B), we could be observing a skew in the results.
+Another aspect that
 

### 2024-11-03 scripts/graph_modelling/step_up_by_state.py

@@ -59,7 +59,6 @@ plt.axhline(270, color='purple', linestyle='--', linewidth=2)
 plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
 
 # Add lines for current Democratic and Republican thresholds
-plt.axhline(225, color='blue', linestyle='--', linewidth=1.5)
 plt.text(210, 220, '225 votes', horizontalalignment='center', color='blue', fontsize=12)
 
 # Add Labels and Title

### 2024-11-03 paper/paper.qmd
Discussion section addtions

@@
 
-plt.figure(figsize=(18, 14))
+plt.figure(figsize=(20, 14))
 plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
@@
 
+The result of the model tells us that overall the polls in october are favouringh Harris to win the election.
+From th perpctiv of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more possible to flip than others.
+North Carolina is the premier state predicted to go blue with the highest liklyhood of Trump winning.
+It would be recommended to allocate resources here to try and win a state from Harris.
+As far as retaining currently favoured states, Michigan is a weak position for Trump according to the model, nearing a coin flip.
+As this state holds 15 lectoral college votes, 4th most of the swing states, it is a key state to win.
+Assuming the model's correct prediction that Michigan is currently held by Trump, expect the democrats to attempt to tip the scales in their favor here.
+
 One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets.
@@
 However, this does not seem to be happening, which suggests that the markets may be pricing in information that the polls do not capture—perhaps reflecting shifts in voter sentiment, hidden preferences, or systematic biases in polling.
-
-## What Else Do We Learn About the World?
 
@@
 In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day.
-Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support [@clinton2021task, @kennedy2018evaluation].
+Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support (@clinton2021task , @kennedy2018evaluation).
 Should this trend continue, and be present in the polls used in this analysis, it may be expected that Trump could surprise the model and win some of these swing states.

### 2024-11-03 paper/paper.qmd
discussion addition and paper editing

@@
 
-This section will discuss a particular poll from the database of polls used in the above analysis, one made by sienna and the new york times.
+This section will discuss a particular poll from the database of polls used in the above analysis, one made by Siena College and the new york times.
 The New York Times partners with the Siena college in order to carry out political polls throughout the election period, these are often high quality polls according to their rating, likely because the nyt has been doing these polls for a long time.
@@
 This action will focus on one poll done in Philadelphia, but those in other states follow the same methodology and patterns.
-This consistency is important for the NYT when they are reporting their results, as it makes comparison around the US viable and reliable.
+This consistency is important for the NYT when they are reporting their results, as it makes comparison around the US states viable and reliable.
 In going over the methodology this appendix aims to understand what about this pollster makes them high quality and what might they leave wanting.
@@
 The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on it's own more redundant.
-(https://hdsr.mitpress.mit.edu/pub/ejk5yhgv#n0xinrk8s1g).
+(@bailey2023new).
 This is clearly a factor here as the response rate was below 2%.
-The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues [@Brick2013Nonresponse].
-The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population.
-Making the means generate more accurate.
+The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse).
+The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generate more accurate.
 With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias.
 It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias.
-Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 [Clinton t al (2021) kennedy et al (2018)].
+Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation).
 With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.
@@
 For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are.
-One documented phenomenon worth mentioning is social desirability bias, that people do tend to choose to answer {definition] (cite).
-This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote ().
-The pollster’s use of both modelled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically.
+One documented phenomenon worth mentioning is social desirability bias, that people do are more inclined to answer with more socially acceptable answers.
+This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@social).
+The pollster’s use of both modeled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically.
 The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however siena/NYT took a methodology that aims to negate this as much as possible.
 Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides.
-Qualitative insights to voters' choices or indecision creates a more robust and significant dataset.
+Qualitative insights to voter's choices or indecision creates a more robust and significant dataset.
 By comparison this is far more valuable than a singular binary choice quiz you may see online.

### 2024-11-03 paper/paper.qmd
Appendix B referencing

@@
 About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish.
-In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (https://www.nytimes.com/interactive/2024/09/19/us/politics/times-siena-inquirer-poll-pennsylvania-likely-electorate.html).
+In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (@nyt2024cross_tabs).
 This high call volume is key to maximising response from all groups.
@@
 One documented phenomenon worth mentioning is social desirability bias, that people do are more inclined to answer with more socially acceptable answers.
-This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@social).
+This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@socialdesire).
 The pollster’s use of both modeled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically.

### 2024-11-03 paper/paper.qmd
my edits

--- 2024-11-03 paper/paper.qmd
@@
 
-This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likly decide th elections result.
-By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump's victory in each state.
-This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.
+This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris.
+By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump or Harris victory in each of the seven swing states.
+This approach allows us to analyze voter support patterns across the states and predict election outcomes more accurately.
 
-The estimand in this study is the probability that Donald Trump wins a given state, derived from aggregated state-level polling averages.
+The estimand in this study is the probability that Donald Trump or Kamala Harris wins a given state, derived from aggregated state-level polling averages.
 The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, with predictor variables comprising the weighted average polling percentages for both candidates.
@@
 The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...
-The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
-The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...
 
@@
 This dataset provides predictions for candidate support for the 2024 U.S.
-Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
-the surveys utilized one or multiple of the following method for rsponse gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
-The diversity of the methods of collection supplies for a colorful array of polls capturing all facets of voters.
+Presidential Election by aggregating various polls that capture public sentiment toward candidates Donald Trump and Kamala Harris.
+Each entry in the dataset corresponds to a specific question about voting preference in a poll conducted by different polling organizations, which measures the percentage of respondents expressing their support for each candidate in their respective state.
+The surveys utilized one or multiple of the following methods for response gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
+The diversity of the methods of collection supplies a colourful array of polls capturing all facets of voters.
 
@@
 
-A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
-States that do not fall under this definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
+A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President" (NEEDS CITATION) .
+States that do not fall under his definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
 Thus the election result is normally decided by which party wins just a handful of states in the electoral college system.
 By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
-Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin; dtermined to be swing states in 2024 [@nyt_swing, @FiveThirtyEight].
-In figure 1 we see the current distribution of electoral votes that each party hold according to those sources as of October 20th, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
+Thus our analysis restricts the datasets to state-level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin [@nyt_swing, @FiveThirtyEight].
+In figure 3 we see the amount that each party hold according to those sources, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump receives over 270th electoral votes.
 
 ```{python}
-#| ccho: false
-#| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, accordng to 538.
-The chart distinguishes between Democrat, Republican, and swing states, with a color-coded fill under each party’s cumulative curve.
-Key threshold of 270 electoral vote mark needed to secure a victory is diplayed.
-This visualization shows the democrats lead with 227 votes and the Republicans with 219 vot secured"
 # Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot
-
 states = ["California", "Connecticut", "Delaware", "Hawaii", "Illinois", "Maine", 
@@
 
-plt.figure(figsize=(20, 14))
+plt.figure(figsize=(18, 8))
 plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
@@
 plt.fill_between(x[:dem_threshold + 1], cumulative_votes[:dem_threshold + 1], step='post', color='blue', alpha=0.4)
-threshold_index = np.argmax(cumulative_votes >= 219)
-plt.fill_between(x_reversed[:threshold_index + 5], cumulative_votes_reversed[:threshold_index + 5], step='post', color='red', alpha=0.4)
 
 # Adding threshold line and markers
-plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
 plt.axhline(270, color='purple', linestyle='--', linewidth=2)
@@
 plt.ylabel('Total Votes', fontsize=12)
-plt.title('Cumulative Electoral Votes by State', fontsize=25)
+plt.title('Cumulative Electoral Votes by State', fontsize=15)
 plt.grid(True, linestyle='--', alpha=0.5)
@@
 Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
-a good score reflecting low propensity to be impacted by errors and biases that can occur in survey sampling.
-It also accounts for impacts that might bias a survey's score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
-Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
-By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
+A good score reflects a low propensity to be impacted by errors and biases that can occur in survey sampling.
+It also accounts for impacts that might bias a survey's score, such as adjustments for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
+Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question-wording, sampling methods, and weighting procedures.
+By combining these two metrics into a single-star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
 We use this single measurement out of 3 to ensure the integrity of our analysis.
@@
 In keeping with this aim, we focused our analysis on polls conducted after September.
-Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, for example Biden dropping out as a candidate on July 21st.
+Polls can be significantly influenced by initial campaign events early in an election cycle, such as the shocks from the first debates or unexpected changes in the candidate lineup.
+For example, Biden dropping out as a candidate on July 21st.
 These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
@@
 ```{python}
+#| echo: false
 
-#| echo: false
-#| fig-cap: "Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states."
-
-df = pd.read_parquet('../data/02-analysis_data/merged_swing_state_data.parquet')
+# Load the data (replace with your CSV file path)
+df = pd.read_csv('../data/02-analysis_data/merged_swing_state_data.csv')
 
@@
 
-# Rename elements
+# Split 'variable_stat' into 'variable' and 'stat'
 summary_long[['variable', 'stat']] = summary_long['variable_stat'].str.rsplit('_', n=1, expand=True)
-summary_long['variable'] = summary_long['variable'].str.replace('_', ' ', regex=False).str.title()
 
@@
 # Print the table using tabulate
-print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='simple_grid'))
+print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='fancy_grid'))
 ```
-**Table 1: Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states.**
 
@@
 Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data.
-The percentage of respondant favouring a candidate become the key variables for prediction.
-Within the collection of polls in our sample the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump , and `harris_pct` respectively.
-Polls ar taking towards the build up of an election and thus ar taken by repondants at different times.
-Figur 2 shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
-Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
+The percentage of respondents favouring a candidate becomes the key variable for prediction.
+Within the collection of polls in our sample, the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump, and `harris_pct` respectively.
+Polls are taken towards the build-up of an election and thus are taken by respondents at different times.
+Graph x shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
+Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondents will be voting on the day.
 ```{python}
 #| echo: false
-#| fig-cap: "A histogram of the poll quantities, aggregated by the end date of each poll, shows the distribution of polling data over time.
-This chart reveals the frequency of polls leading up to the election and provides insight into polling activity trends, with potential spikes reflecting periods of increased public interest or campaign activity."
 
@@
 plt.ylabel('Poll Count')
-plt.title('Poll Distribution Over Time', fontsize=25)
+plt.title('Poll Distribution Over Time')
 plt.xticks(rotation=45)
@@
 
-Another aspects of the polls we wish to consider is the population size of each.
+Another aspect of the polls we wish to consider is the population size of each.
 Larger polls result in a stronger reduction in randomness in the response.
 The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers.
-In table 1 we can see the variety within the states polls.
-For use in the model the percentages in these polls is weighted for recenecy and size of the population.
+In Table X we can see the variation of sample size within the polls.
+For use in the model, the percentages in these polls are weighted for recency and size of the sample population.
 
@@
 
-```{python}
+ ```{Python}
 #| echo: false
+
+import pandas as pd
+from tabulate import tabulate
+
 # Data setup
@@
     'State': ['Arizona', 'Georgia', 'Michigan', 'Nevada', 'North Carolina', 'Pennsylvania', 'Wisconsin'],
-    'Trump %': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
-    'Trump SD': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
-    'Harris %': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
-    'Harris SD': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
-    'Trump Winning %': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
+    'Trump % (Mean)': [49.85, 49.39, 48.57, 47.9, 48.48, 47.68, 47.76],
+    'Trump % (SD)': [1.24, 1.78, 2.14, 1.5, 1.41, 1.81, 2.0],
+    'Harris % (Mean)': [47.04, 47.66, 47.72, 48.63, 48.58, 48.53, 48.92],
+    'Harris % (SD)': [2.17, 2.52, 2.12, 1.19, 1.55, 1.29, 1.72],
+    'Percentage Chance of Trump Winning': [86.85, 71.18, 53.09, 21.34, 32.31, 19.67, 15.52]
 }
@@
 # Display as a professional table
-print(tabulate(
+table_output = tabulate(
     df, headers='keys', tablefmt='grid', showindex=False,
     floatfmt=(".2f", ".2f", ".2f", ".2f", ".2f", ".2f")
-))
+)
+print(table_output)
 
 ```
-**Table 2: Results table for the model describd above.
-Show support fo trump and Kamala at the stat level accordng to avalable polls.
-Small standard dviation show strong results, yet the rac is very close in many states.
-Percentages above 50% show the favoured candidate in each state.**
+
+in figure xx we can see the results from the model described.
 
 In Arizona and Georgia, the model predicts the strongest likelihood of a Trump victory, with a mean polling percentage showing Trump over 2 percentage points ahead in Arizona.
-In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.The data is thus resulting in an strong 86.85% and and 71.18% likelyhood in Trump winning in Arizona and Georgia respectively.
-Trump is also predicted to win in Michigan, although by smaller margins as shown in this state.
-As a result the model predicts trump winning with only a 53% chance, showing his odds are near to a coin flip.
-With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.
+In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.
+The data thus results in a strong 86.85% and 71.18% likelihood of Trump winning in Arizona and Georgia respectively.
+The model predicts Trump has a 53% chance of winning Michigan. With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.
 
@@
 #| echo: false
-#| fig-cap: "A centered bar chart where each state has a bar reflecting Trump’s probability of winning.
-The bars extend to the right or left from the center line at 50%, with the distance representing the magnitude of deviation from an even chance.
-This chart visualizes Trump’s relative strength in each swing state and highlights that he s only favoured in Georgia and Arizona."
+
 # Data setup (can adjust based on model results)
@@
 ax.set_xlabel('Distance from 50% (Trump Winning Probability)')
-ax.set_title('Probability of Trump Win in swing states, Centered from 50%', fontsize=25)
+ax.set_title('Probability of Trump Win, Centered from 50%')
 plt.tight_layout()
@@
 
-Kamala is predicted to win the remaining states and is given fairly high chances of locking these up.
-despite only winning Nevada, North Carolina and Pennsilvania by under a percntage point in the model Kamala's probablity of winning those electoral votes is very high.
-This is due to the small standard deviations shown in table 2 making the predicted outcome of a victory more secure.
-Wisconson is shown to be her srongest and most likley victory of the swing states, with Trump only having a 15% chance here. 
-Overall this givs Harris 51 of the 93 swing state votes, our model expecting a win for the Democrats and the first female US president!
+Kamala is predicted to win the remaining of the states and is given fairly high chances of locking these up.
+despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model Kamala's predicted probability of winning those electoral votes is very high.
+This is due to the small standard deviations shown in Table X making the predicted rule of the states leaning Democrat being more likely.
+Wisconsin is shown to be her strongest and most likely victory of the swing states. 
+Overall this gives Harris 51 of the 93 swing state votes, this will result in a win for the democrats and the first female president of the United States!
 
 ## Trump overall probability
-Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and loses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes.
-Based on the modeled probabilities his current expected votes is 41.05, below the threshold of 51 he needs based on our assumptions.
+-- Taking the probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and losses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes.
+Based on the model probabilities his current expected vote is 41.05, below the threshold of 51 he needs based on our assumptions.
 
-Seperating the outcomes into those that give trump the required electoral college votes, we can obtain probability that trump seizes a second term.
-Using probability theory, and assuming independence among the states outomes, in our approach.
+Separating the outcomes into those that give Trump the required electoral college votes, we can obtain the probability that Trump seizes a second term.
+
+Using a probability theory, and assuming independence among the states, in our approach.
 When accounting for all possible state outcome combinations, the model calculates that Trump has approximately a 22.5% chance of securing at least 51 electoral votes from these key swing states.
@@
 
-The result of the model tells us that overall the polls in october are favouringh Harris to win the election.
-From th perpctiv of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more possible to flip than others.
-North Carolina is the premier state predicted to go blue with the highest liklyhood of Trump winning.
-It would be recommended to allocate resources here to try and win a state from Harris.
-As far as retaining currently favoured states, Michigan is a weak position for Trump according to the model, nearing a coin flip.
-As this state holds 15 lectoral college votes, 4th most of the swing states, it is a key state to win.
-Assuming the model's correct prediction that Michigan is currently held by Trump, expect the democrats to attempt to tip the scales in their favor here.
+Without the election results to validate our model, it is difficult to extrapolate much meaning from this predictive model.
+However, we do learn that when this polling data was taken, Kamala Harris was leading the polls in the swing states.
+As we get closer to the election, the polls have narrowed and show a lean in Trump's direction.
+Having updated the data closer to the election would likely result in a different model outcome.
+This teaches us that the proximity to the election may be of particular importance with regard to predictive power, or that it is possibly less useful and maybe something that should be weighed less heavily in our model.
+
+## What Else Do We Learn About the World?
 
 One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets.
-While the majority of reputable polls suggest that Kamala Harris is favored on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
+While the majority of reputable polls suggest that Kamala Harris is favoured on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
 This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.
@@
 Another important consideration is the possible biases in both polling and betting markets.
-Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favor of Harris.
-Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favor him.
+Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favour of Harris.
+Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favour him.
 Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities.
@@
 First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals).
-As polling results are based on public opinion, large shock events are likely and expected, moreover scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modeled here.
-Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve nonlinear influences that are difficult to predict.
+Additionally, the use of linear modelling may oversimplify the complex dynamics of voter behaviour, as elections often involve non-linear influences that are difficult to predict.
 
@@
 Training the model on a more comprehensive dataset that includes these states and other predictive variables, such as demographic factors or economic indicators, could lead to a more accurate forecast.
-
-Candidate Trump’s low likelihood of winning shown by our results may under-represent his support, this claim is supported by the disparity with the betting markets.
-This could be a result of a common problem with polling data, response rate bias.
-In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day.
-Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support (@clinton2021task , @kennedy2018evaluation).
-Should this trend continue, and be present in the polls used in this analysis, it may be expected that Trump could surprise the model and win some of these swing states.
-Another reason Trump could outperform the model is another bias often seen in surveys, one of social desirability.
-Trump is a unique candidate as, if elected, he would become the first president to be elected with federal crime conviction.
-This along with other negative publicity the former host of tv’s apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support.
-Even with the quality polls used in this analysis, where steps are taken to ensure representative results (see appendix B), we could be observing a skew in the results.
-Another aspect that
 
@@
 Presidential Election, there is still much to learn.
-By refining the model and incorporating additional data, we can move closer to producing predictions that more accurately reflect voter behavior and election outcomes.
+By refining the model and incorporating additional data, we can move closer to producing predictions that more accurately reflect voter behaviour and election outcomes.
 
@@
 Each participant will receive a $10 allowance to bet on their predicted winner (Trump or Harris), with payouts based on live odds, creating a direct incentive for respondents to make thoughtful, informed predictions.
-The purpose of this is twofold, it aims to minmize the attrition of the respondants, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win.
-Ideally, the purpose of this is so that people take into consideration who their friends and family, as well as their community are expecting to win.
+The purpose of this is twofold, it aims to minimize the attrition of the respondents, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win.
+Ideally, the purpose of this is so that people take into consideration who their friends and family, as well as their community, are expecting to win.
 
@@
 
-This section will discuss a particular poll from the database of polls used in the above analysis, one made by sienna and the new york times.
-The New York Times partners with the Siena college in order to carry out political polls throughout the election period, these are often high quality polls according to their rating, likely because the nyt has been doing these polls for a long time.
-The NYT polls often emphasise a focus on the battleground states and poll then more than others making these very useful in this paper’s analysis.
+This section will discuss a particular poll from the database of polls used in the above analysis, one made by Siena and the New York Times.
+The New York Times partners with the Siena College to carry out political polls throughout the election period, these are high-quality polls according to their rating, likely because the NYT has been doing these polls for a long time.
+The NYT polls often emphasize a focus on the battleground states and poll them more than others making these very useful in this paper’s analysis.
 This action will focus on one poll done in Philadelphia, but those in other states follow the same methodology and patterns.
-This consistency is important for the NYT when they are reporting their results, as it makes comparison around the US viable and reliable.
+This consistency is important for the NYT when they are reporting their results, as it makes comparisons around the US viable and reliable.
 In going over the methodology this appendix aims to understand what about this pollster makes them high quality and what might they leave wanting.
@@
 The poll employs a response rate-adjusted stratified sampling method using the L2 voter file, which contains detailed demographic information on registered voters.
-Initially the interviewee's are selected randomly from a national list of registered voters, response are then weighted.
+Initially, the interviewees are selected randomly from a national list of registered voters, responses are then weighted.
 The stratification, often referred to as weighing, is applied to the random sample collected in order to properly reflect the entire demographic.
@@
 - Age
-- Home ownership
-The sampling is conducted in order to address telephone coverage around the state as well as nonresponse rates.
-Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Phyladelphia,.
-The voter file is then ratified and weights are based on historically modelled response ra and coverage.
+- Homeownership
+The sampling is conducted to address telephone coverage around the state as well as nonresponse rates.
+Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Philadelphia.
+The voter file is then ratified and weights are based on historically modelled response rate and coverage.
 
@@
 Times/ Siena polls are conducted by telephone, using live interviews.
-About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish.
+About 96% of interviews were conducted by cellphone, and interviews were conducted in both English and Spanish.
 In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (https://www.nytimes.com/interactive/2024/09/19/us/politics/times-siena-inquirer-poll-pennsylvania-likely-electorate.html).
-This high call volume is key to maximising response from all groups.
-In a further effort to get a good sample additional effort is made to constant under-represented groups, for example people with lower education levels.
+This high call volume is key to maximizing response from all groups.
+In a further effort to get a good sample additional effort is made to constantly under-represented groups, for example, people with lower education levels.
 
@@
 The questions themselves included more than just the information on who a participant would vote on.
-Followup questions were given either because or regardless of what their previous answers were.
-One example would be follow up questions for those who answer as “don’t know/ refused” in order to understand some reasoning behind this answer.
+Follow-up questions were given either because or regardless of what their previous answers were.
+One example would be follow-up questions for those who answer as “don’t know/ refused” in order to understand some reasoning behind this answer.
 Those who did select an option also were asked questions to get a clear image of their sentiment or reasoning.
@@
 Qualitative insights to voters' choices or indecision creates a more robust and significant dataset.
-By comparison this is far more valuable than a singular binary choice quiz you may see online.
+By comparison, this is far more valuable than a singular binary choice quiz you may see online.
 
@@
 This poll’s methodology makes great efforts to mitigate these biases through its weighting methods, yet these factors could affect the accuracy of the poll results.
-Nonetheless, The poll provides valuable insights and remains a useful resource for analysing voter preferences, especially due to its aforementioned weighting strategy.
+Nonetheless, The poll provides valuable insights and remains a useful resource for analyzing voter preferences, especially due to its aforementioned weighting strategy.
 The poll's transparency in methodology makes it an even more reliable resource and enhances the credibility and comparability of the findings.

### 2024-11-03 paper/paper.qmd
add abstract

--- 2024-11-03 paper/paper.qmd
@@
 abstract: |
-
- This paper is about the election
+ "This paper develops a predictive model for the 2024 U.S.
+Presidential Election, analyzing state-level polling data to forecast outcomes between Donald Trump and Kamala Harris.
+Using logistic regression, the model estimates the likelihood of Trump’s success in swing states by weighting polls based on recency and sample size.
+Findings indicate that Harris has a higher probability of winning in most swing states, though close margins suggest targeted campaign efforts could alter outcomes in pivotal areas.
+This analysis provides insight into electoral dynamics, highlighting how data-driven approaches can clarify potential election outcomes and inform strategic political decision-making."
 ---
@@
 In going over the methodology this appendix aims to understand what about this pollster makes them high quality and what might they leave wanting.
+
 ## Sampling method
+
 The poll employs a response rate-adjusted stratified sampling method using the L2 voter file, which contains detailed demographic information on registered voters.
@@
 The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generate more accurate.
+
 With such low response rates the possibility for risk, in spite of the effort made by the pollster, is that non-response and choice are correlated, this is known as non-response bias.
@@
 With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.
+
 The benefit of how the survey was taken is the detail added from the additional questions asked, some key questions allow for further stratification based on the answers.
@@
 
-## what is included in the poll?
-
-## Methodology and why?
-
 # References

### 2024-11-03 paper/paper.qmd
some updates

@@
 abstract: |
- "This paper develops a predictive model for the 2024 U.S.
+ This paper develops a predictive model for the 2024 U.S.
 Presidential Election, analyzing state-level polling data to forecast outcomes between Donald Trump and Kamala Harris.
@@
 Findings indicate that Harris has a higher probability of winning in most swing states, though close margins suggest targeted campaign efforts could alter outcomes in pivotal areas.
-This analysis provides insight into electoral dynamics, highlighting how data-driven approaches can clarify potential election outcomes and inform strategic political decision-making."
+This analysis provides insight into electoral dynamics, highlighting how data-driven approaches can clarify potential election outcomes and inform strategic political decision-making.
 ---
@@
 
-# 1 Introduction
-
-This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likly decide th elections result.
-By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump's victory in each state.
+# Introduction
+This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likely decide the election result.
+By aggregating high-quality polls that account for recency and sample size, we create a logistic regression model that estimates the probability of a Trump victory in each state.
 This approach allows us to analyze voter support patterns across the states and predict election outcomes from polling data.
 
-The estimand in this study is the probability that Donald Trump wins a given state, derived from aggregated state-level polling averages.
+The estimand in this study is the probability that Donald Trump will win a given state, derived from aggregated state-level polling averages.
 The binary outcome variable in our model indicates whether Trump (1) or Harris (0) is predicted to win in each state, with predictor variables comprising the weighted average polling percentages for both candidates.
@@
 We have focused our analysis on the swing states where polling percentages are closely contested and could significantly influence the final election result.
-
 Accurate election predictions provide valuable insights into voter dynamics, helping political analysts, campaigns, and the public anticipate electoral outcomes.
@@
 # Data {#sec-data}
-
 The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
-The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...
-The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
-The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], Sklearn [@Sklearn]...
+The programming language for data analysis, visualization and statistical investigation was Python [@Python] along with the packages Matplot [@Matplot], Seaborn [@Seaborn], Numpy [@Numpy], Pandas [@Pandas], Jupyter [@Jupyter], Tabulate [@Tabulate], and Sklearn [@Sklearn].
 
 # 2.1 Measurement
-
 The voter support data used in this analysis comes from raw polling information sourced from the Project 538 online database [@FiveThirtyEight].
 This dataset provides predictions for candidate support for the 2024 U.S.
-Presidential Election by aggregating various polls that capture public sentiment towards candidates Donald Trump and Kamala Harris.
-Each entry in the dataset corresponds to a specific question around voting preference in a poll conducted by different polling organisations, which measure the percentage of respondents expressing their support for each candidate in their respective state.
-the surveys utilized one or multiple of the following method for rsponse gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
-The diversity of the methods of collection supplies for a colorful array of polls capturing all facets of voters.
+Presidential Election by aggregating various polls that capture public sentiment toward candidates Donald Trump and Kamala Harris.
+Each entry in the dataset corresponds to a specific question about voting preference in a poll conducted by different polling organizations, which measures the percentage of respondents expressing their support for each candidate in their respective state.
+the surveys utilized one or multiple of the following methods for response gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
+The diversity of the methods of collection supplies a colourful array of polls capturing all facets of voters.
 
@@
 # 2.2 The Dataset
-
 A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
@@
 By concentrating on these swing states, we aim to capture the most critical and uncertain areas of the electoral map, where voter preferences are most likely to sway the final result.
-Thus our analysis restricts the datasets to state level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin; dtermined to be swing states in 2024 [@nyt_swing, @FiveThirtyEight].
-In figure 1 we see the current distribution of electoral votes that each party hold according to those sources as of October 20th, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump recieve over 270th electoral votes.
+Thus our analysis restricts the datasets to state-level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin; determined to be swing states in 2024 [@nyt_swing, @FiveThirtyEight].
+In @Fig-1 we see the current distribution of electoral votes that each party hold according to those sources as of October 20th, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump receives over 270th electoral votes.
 
@@
 #| ccho: false
+#| label: Fig-1
 #| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, accordng to 538.
@@
 
-It’s important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
+It is important to note that each pollster employs unique sampling techniques and methodologies that influence their Pollscore and Transparency Score.
 Pollscore is a measure of a pollster's historical accuracy, reflecting the quality of their sampling methods and data collection.
-a good score reflecting low propensity to be impacted by errors and biases that can occur in survey sampling.
-It also accounts for impacts that might bias a survey's score, such as adjustmenting for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
-Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question wording, sampling methods, and weighting procedures.
-By combining these two metrics into a single star rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
+A good score reflects low propensity to be impacted by errors and biases in survey sampling.
+It also accounts for impacts that might bias a survey's score, such as adjustments for the differing difficulty of polling certain political races or elements of luck in samples by resampling polls.
+Transparency Score, on the other hand, measures how openly a pollster shares methodological details like question-wording, sampling methods, and weighting procedures.
+By combining these two metrics into a rating, we achieve an excellent overall assessment of a pollster's quality, capturing both their empirical accuracy and their commitment to methodological transparency.
+
 We use this single measurement out of 3 to ensure the integrity of our analysis.
@@
 In keeping with this aim, we focused our analysis on polls conducted after September.
-Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup, for example Biden dropping out as a candidate on July 21st.
+Early in the election cycle, polls can be significantly influenced by initial campaign events, such as the shocks from the first debates or unexpected changes in the candidate lineup.
+For example, Joe Biden dropping out as a candidate on July 21st severely impacted the odds of a Kamala Harris victory.
 These factors can introduce biases and fluctuations that do not necessarily reflect the enduring preferences of the electorate.
@@
 #| echo: false
+#| label: Fig-2
 #| fig-cap: "Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states."
@@
 Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data.
-The percentage of respondant favouring a candidate become the key variables for prediction.
-Within the collection of polls in our sample the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump , and `harris_pct` respectively.
-Polls ar taking towards the build up of an election and thus ar taken by repondants at different times.
-Figur 2 shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
-Because sentiment changes over time we will adjust the weight that we place on polls that are closer to the election as these will be more indicative of how respondants will be voting on the day.
+The percentage of respondents favouring a candidate is the key variable used for prediction.
+Within the collection of polls in our sample, the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump, and `harris_pct` respectively.
+
+Polls are taken towards the build-up of an election and thus are taken by respondents at different times.
+@Fig-2 shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
+Because sentiment changes over time we adjust the weight that we place on polls that are closer to the election as these are more indicative of how respondents will be voting on the day.
+
 ```{python}
 #| echo: false
+#| label: Fig-3
 #| fig-cap: "A histogram of the poll quantities, aggregated by the end date of each poll, shows the distribution of polling data over time.
@@
 #| echo: false
+#| label: Fig-4
 # Data setup
@@
 #| echo: false
+#| label: Fig-5
 #| fig-cap: "A centered bar chart where each state has a bar reflecting Trump’s probability of winning.

### 2024-11-03 paper/paper.qmd
reformat/update paper

@@
 
-# Introduction
+# 1 Introduction
 This paper develops a predictive model for the 2024 US Presidential Election using state-level polling data to forecast the likely winner between Donald Trump and Kamala Harris, focusing on swing states that will likely decide the election result.
@@
 
-# Data {#sec-data}
+# 2 Data {#sec-data}
 The data used in this paper was gathered from FiveThirtyEight [@FiveThirtyEight] a website that aims to use "data and evidence to advance public knowledge".
@@
 ```
-**Table 1: Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states.**
+**Figure 2: Summary statistics of key sample variables by state. Shows high sample size and numeric grade among all polls of swing states.**
 
@@
 
-Another aspects of the polls we wish to consider is the population size of each.
+Another aspect of the polls we wish to consider is the population size of each.
 Larger polls result in a stronger reduction in randomness in the response.
 The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers.
-In table 1 we can see the variety within the states polls.
-For use in the model the percentages in these polls is weighted for recenecy and size of the population.
+In @Fig-3 we can see the variety within the state polls.
+For use in the model, the percentages in these polls are weighted for recency and size of the sampled population.
 
@@
 ```
-**Table 2: Results table for the model describd above.
-Show support fo trump and Kamala at the stat level accordng to avalable polls.
-Small standard dviation show strong results, yet the rac is very close in many states.
+**Figure 4: Is a results table for the model described above.It shows support for Trump and Harris at the state level according to the available polls.
+Small standard deviations show consistent results, yet the race is very close in many states.
 Percentages above 50% show the favoured candidate in each state.**
@@
 In Arizona and Georgia, the model predicts the strongest likelihood of a Trump victory, with a mean polling percentage showing Trump over 2 percentage points ahead in Arizona.
-In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.The data is thus resulting in an strong 86.85% and and 71.18% likelyhood in Trump winning in Arizona and Georgia respectively.
+In Georgia, Trump's mean polling percentage is 49.39% against Harris's 47.66%.
+The data thus resulted in a strong 86.85% and 71.18% likelihood of Trump winning in Arizona and Georgia respectively.
+
 Trump is also predicted to win in Michigan, although by smaller margins as shown in this state.
-As a result the model predicts trump winning with only a 53% chance, showing his odds are near to a coin flip.
+As a result, the model predicts Trump winning with only a 53% chance, showing his odds are near to a coin flip.
 With Arizona, Georgia and Michigan Trump is predicted a total of 42 of the 93 swing state votes.
@@
 Kamala is predicted to win the remaining states and is given fairly high chances of locking these up.
-despite only winning Nevada, North Carolina and Pennsilvania by under a percntage point in the model Kamala's probablity of winning those electoral votes is very high.
-This is due to the small standard deviations shown in table 2 making the predicted outcome of a victory more secure.
-Wisconson is shown to be her srongest and most likley victory of the swing states, with Trump only having a 15% chance here. 
-Overall this givs Harris 51 of the 93 swing state votes, our model expecting a win for the Democrats and the first female US president!
+despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model Kamala's probability of winning those electoral votes is very high.
+This is due to the small standard deviations shown in @Tab-2 making the predicted outcome of a victory more secure.
+
+Wisconsin is shown to be her strongest and most likely victory of the swing states, with Trump only having a 15% chance here.
+
+Overall this gives Harris 51 of the 93 swing state votes, our model expecting a win for the Democrats and the first female US president!
 
 ## Trump overall probability
-Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and loses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes.
-Based on the modeled probabilities his current expected votes is 41.05, below the threshold of 51 he needs based on our assumptions.
 
-Seperating the outcomes into those that give trump the required electoral college votes, we can obtain probability that trump seizes a second term.
-Using probability theory, and assuming independence among the states outomes, in our approach.
+Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and losses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes.
+Based on the modelled probabilities his current expected vote is 41.05, below the threshold of 51 he needs based on our assumptions.
+
+By separating the outcomes into those that give Trump the required electoral college votes, we can obtain the probability that Trump seizes a second term, using probability theory, and assuming independence among the state outcomes in our approach.
 When accounting for all possible state outcome combinations, the model calculates that Trump has approximately a 22.5% chance of securing at least 51 electoral votes from these key swing states.
@@
 
-The result of the model tells us that overall the polls in october are favouringh Harris to win the election.
-From th perpctiv of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more possible to flip than others.
-North Carolina is the premier state predicted to go blue with the highest liklyhood of Trump winning.
+The result of the model tells us that overall the polls in October are favouring Harris to win the election.
+From the perspective of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more likely to flip than others.
+North Carolina is the premier state predicted to go blue with the highest likelihood of Trump winning.
 It would be recommended to allocate resources here to try and win a state from Harris.
-As far as retaining currently favoured states, Michigan is a weak position for Trump according to the model, nearing a coin flip.
-As this state holds 15 lectoral college votes, 4th most of the swing states, it is a key state to win.
-Assuming the model's correct prediction that Michigan is currently held by Trump, expect the democrats to attempt to tip the scales in their favor here.
+As far as retaining currently favoured states, Michigan is in a weak position for Trump according to the model, nearing a coin flip.
+As this state holds 15 electoral College votes, the 4th most of the swing states, it is a key state to win.
+Assuming the model's correct prediction that Michigan is currently held by Trump, expect the Democrats to attempt to tip the scales in their favour there.
 
 One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets.
-While the majority of reputable polls suggest that Kamala Harris is favored on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
+While the majority of reputable polls suggest that Kamala Harris is favoured on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
 This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.
@@
 Another important consideration is the possible biases in both polling and betting markets.
-Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favor of Harris.
-Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favor him.
+Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favour of Harris.
+Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favour him.
 Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities.
-However, historical data still suggest that polls have been accurate 78% of the time in predicting elections [@FiveThirtyEight], indicating that the reliability of betting markets remains questionable, especially in light of their poor performance during the 2016 election, when Trump’s odds were listed at +475 (17%) just before his victory [https://www.oddsshark.com/entertainment/us-presidential-odds-2016-futures].
+However, historical data still suggest that polls have been accurate 78% of the time in predicting elections [@FiveThirtyEight], indicating that the reliability of betting markets remains questionable, especially in light of their poor performance during the 2016 election, when Trump’s odds were listed at +475 (17%) just before his victory [@oddsshark2016].
 
@@
 First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals).
-As polling results are based on public opinion, large shock events are likely and expected, moreover scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modeled here.
-Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve nonlinear influences that are difficult to predict.
+As polling results are based on public opinion, large shock events are expected, furthermore, scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modelled here.
+Additionally, the use of linear modelling may oversimplify the complex dynamics of voter behaviour, as elections often involve nonlinear influences that are difficult to predict.
 
@@
 
-Candidate Trump’s low likelihood of winning shown by our results may under-represent his support, this claim is supported by the disparity with the betting markets.
+Trump’s low likelihood of winning shown by our results may under-represent his support, this claim is supported by the disparity with the betting markets.
 This could be a result of a common problem with polling data, response rate bias.
 In the previous two elections, where Donald Trump was also the republican candidate, his polling was far below the eventual results on election day.
-Post-mortems of both elections indicated evidence of those supporting Trump not participating in polls and thus under-presenting Republican support (@clinton2021task , @kennedy2018evaluation).
-Should this trend continue, and be present in the polls used in this analysis, it may be expected that Trump could surprise the model and win some of these swing states.
+Post-mortems of both elections indicated evidence that those supporting Trump were not participating in polls and thus under-presenting Republican support (@clinton2021task , @kennedy2018evaluation).
+Should this trend continue Trump could surprise the model and win some of these swing states.
 Another reason Trump could outperform the model is another bias often seen in surveys, one of social desirability.
-Trump is a unique candidate as, if elected, he would become the first president to be elected with federal crime conviction.
-This along with other negative publicity the former host of tv’s apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support.
-Even with the quality polls used in this analysis, where steps are taken to ensure representative results (see appendix B), we could be observing a skew in the results.
-Another aspect that
+Trump is a unique candidate as, if elected, he would become the first president to be elected with a federal crime conviction.
+This along with other negative publicity the former host of TV’s Apprentice has garnered since the start of his political campaign could bias voters from publicly or privately showing their support.
+Even with the quality polls used in this analysis, where steps are taken to ensure representative results (see Appendix B), we could be observing a skew in the results.
 
@@
 - **Demographics**: Race, age, gender, education, and income.
+
 - **Geography**: Emphasis on swing states with representation from urban, suburban, and rural areas.
+
 - **Voter History**: Including respondents with varying voting patterns, such as frequent voters, occasional voters, and those who are less likely to vote.
@@
 - **Online Recruitment**: Targeted ads on social media, political forums, and news websites will attract a wide audience, highlighting the unique opportunity to place a small bet on the election outcome.
+
 - **Telephone Recruitment**: To capture older demographics and those less reachable online, we will conduct phone outreach using commercial databases and voter registration information.
@@
 ### Incentives for Engagement
+
 A betting pool is a central part of this methodology, with $50,000 allocated as rewards for respondents who accurately predict the election outcome.
 Each participant will receive a $10 allowance to bet on their predicted winner (Trump or Harris), with payouts based on live odds, creating a direct incentive for respondents to make thoughtful, informed predictions.
-The purpose of this is twofold, it aims to minmize the attrition of the respondants, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win.
-Ideally, the purpose of this is so that people take into consideration who their friends and family, as well as their community are expecting to win.
+The purpose of this is twofold, it aims to minimize the attrition of the respondents, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win.
+Ideally, the purpose of this is so that people take into consideration who their friends, family, and their community, are expecting to win.
 
@@
 - **Cross-Referencing Survey Responses**: Survey responses will be cross-referenced with voter registration data to validate eligibility.
+
 - **Weighting**: Data will be weighted to reflect broader population demographics, adjusting for any imbalances in representation across age, race, gender, and geographic factors.
@@
 - **Sample Size and Recency**: More recent polls and those with larger sample sizes will be weighted more heavily.
+
 - **Pollster Reliability**: Historical poll accuracy and transparency scores will further inform the weight of each poll.
@@
 - **Recruitment**: $10,000 for targeting and engaging respondents across multiple platforms.
+
 - **Survey Administration (Online and Phone)**: $20,000 for both online and phone-based survey collection.
+
 - **Betting Pool**: $50,000 allocated to reward accurate predictions and enhance response quality.
+
 - **Data Validation and Analysis**: $10,000 for verification and weighting.
+
 - **Modeling**: $10,000 for analyzing and forecasting based on aggregated data.
 
-## 6. Question Design and Goal Allignment
+## 6. Question Design and Goal Alignment
 The goal of our survey is to predict who will win the election, this is the research question that guides our survey.
-In order to achieve this, our questions will be isolate the relevant variables by asking only one thing at a time (ceteris paribus).
-We will also use item specific scales over agree-disagree formats to avoid acquiescence bias.
-Furthermore, we will randomize response options in order to mitigate response order bias, and emphasize the surveys anonymity to minimize social desireablity bias.
+In order to achieve this, our questions will isolate the relevant variables by asking only one thing at a time (ceteris paribus).
+We will also use item-specific scales over agree-disagree formats to avoid acquiescence bias.
+Furthermore, we will randomize response options in order to mitigate response order bias and emphasize the survey's anonymity to minimize social desirability bias.
 
@@
 
-# Appendix B: Siena / New York Tims Pennsilvania poll
+# Appendix B: Siena / New York Tims Pennsylvania poll
 
-This section will discuss a particular poll from the database of polls used in the above analysis, one made by Siena College and the new york times.
-The New York Times partners with the Siena college in order to carry out political polls throughout the election period, these are often high quality polls according to their rating, likely because the nyt has been doing these polls for a long time.
-The NYT polls often emphasise a focus on the battleground states and poll then more than others making these very useful in this paper’s analysis.
-This action will focus on one poll done in Philadelphia, but those in other states follow the same methodology and patterns.
-This consistency is important for the NYT when they are reporting their results, as it makes comparison around the US states viable and reliable.
-In going over the methodology this appendix aims to understand what about this pollster makes them high quality and what might they leave wanting.
+This section discusses the Siena College and the New York Times poll from the database of polls used in the above analysis.
+The New York Times (NYT) polls often emphasize the swing states making these very useful in this paper’s analysis.
+This section will focus specifically on one poll done in Philadelphia, but those in other states follow the same methodology and patterns.
+This consistency is important for the New York Times when they are reporting their results as it makes comparisons of states reliable.
+In going over the methodology this appendix aims to understand what about this pollster makes them high quality and where they might improve.
 
@@
 The poll employs a response rate-adjusted stratified sampling method using the L2 voter file, which contains detailed demographic information on registered voters.
-Initially the interviewee's are selected randomly from a national list of registered voters, response are then weighted.
+Initially, the interviewees are selected randomly from a national list of registered voters, responses are then weighed.
 The stratification, often referred to as weighing, is applied to the random sample collected in order to properly reflect the entire demographic.
 The stratification accounts for:
+
 - Statehouse district
@@
 The sampling is conducted in order to address telephone coverage around the state as well as nonresponse rates.
-Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Phyladelphia,.
-The voter file is then ratified and weights are based on historically modelled response ra and coverage.
+Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Philadelphia.
+The voter file is then ratified and weights are based on historically modelled response rate and coverage.
 
 ## Collection
-Times/ Siena polls are conducted by telephone, using live interviews.
-About 96% of interviews were contacted by cellphone, and interviews were conducted in both English and spanish.
+Times/Siena polls are conducted by telephone, using live interviews.
+About 96% of interviewees were contacted by cellphone, and interviews were conducted in both English and Spanish.
 In the Philadelphia polls occurring between September 11th and 16th, 240,000 calls were made to nearly 118,000 likely voters (@nyt2024cross_tabs).
-This high call volume is key to maximising response from all groups.
-In a further effort to get a good sample additional effort is made to constant under-represented groups, for example people with lower education levels.
+This high call volume is key to maximizing response from all groups.
+In a further effort to get a good sample additional effort is made to contact under-represented groups.
 
 ## Questionnaire design
+
 The questions themselves included more than just the information on who a participant would vote on.
 Followup questions were given either because or regardless of what their previous answers were.
-One example would be follow up questions for those who answer as “don’t know/ refused” in order to understand some reasoning behind this answer.
+One example would be follow-up questions for those who answer “don’t know/ refused” in order to understand some reasoning behind this answer.
 Those who did select an option also were asked questions to get a clear image of their sentiment or reasoning.
@@
 ## Methodology pros and cons
-The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on it's own more redundant.
+
+The Random contact method in which people are contacted does not provide random sampling, as non-response has made this method on its own more redundant.
 (@bailey2023new).
@@
 The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse).
-The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generate more accurate.
+The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generated more accurate.
 
@@
 It is not uncommon to assume that non-response should occur at random, especially as this poll documents repeatedly calling each likely voter, however in the case of elections there is previous evidence of a bias.
-Research into the previous two elections, and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation).
-With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the Discussion section.
+Research into the previous two elections and the polls proceeding then showed a high probability that non-responders were more likely to vote for Donald Trump, who was the Republican candidate in both 2016 and 2020 (@clinton2021task, @kennedy2018evaluation).
+With Trump running again in 2024, his popularity being underestimated is a distinct possibility, as is mentioned in the discussion section.
 
@@
 For example, survey information is gathered on self-reported likelihood to vote, and obviously this gives respondents answers far more weight if they are.
-One documented phenomenon worth mentioning is social desirability bias, that people do are more inclined to answer with more socially acceptable answers.
+One documented phenomenon worth mentioning is social desirability bias, that people are more inclined to answer with more socially acceptable answers.
 This could occur throughout the questionnaire, however it is very likely that respondents overestimate their likelihood to vote (@socialdesire).
-The pollster’s use of both modeled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically.
-The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however siena/NYT took a methodology that aims to negate this as much as possible.
+The pollster’s use of both modelled turnout probabilities based on historical data and self-reporting improves the accuracy of predicting who is likely to vote dramatically.
+The poll’s use of likely voters still brings some uncertainty to its results, as the results may be impacted by non-voters, however, Siena/NYT took a methodology that aims to negate this as much as possible.
 Another benefit of live interviews is that clarification of answers is possible and the additional depth of the data that this provides.
-Qualitative insights to voter's choices or indecision creates a more robust and significant dataset.
-By comparison this is far more valuable than a singular binary choice quiz you may see online.
+Qualitative insights into voter's choices or indecision create a more robust and significant dataset.
+By comparison, this is far more valuable than a singular binary choice quiz you may see online.
 
@@
 This poll’s methodology makes great efforts to mitigate these biases through its weighting methods, yet these factors could affect the accuracy of the poll results.
-Nonetheless, The poll provides valuable insights and remains a useful resource for analysing voter preferences, especially due to its aforementioned weighting strategy.
+Nonetheless, The poll provides valuable insights and remains a useful resource for analyzing voter preferences, especially due to its aforementioned weighting strategy.
 The poll's transparency in methodology makes it an even more reliable resource and enhances the credibility and comparability of the findings.
 
+## what is included in the poll?
+
+## Methodology and why?
+
 # References

### 2024-11-04 paper/paper.qmd
new title

@@
 ---
-title: "Election 2024"
+title: "Predicting the 2024 U.S. Presidential Election: A Data-Driven Analysis of Swing State Outcomes"
+subtitle: "Forecast Model Suggests Harris as the Probable Winner of the majority of swing states"
 author: "Alexander Guarasci & Jacob Gilbert"
@@
 This is clearly a factor here as the response rate was below 2%.
-The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@Brick2013Nonresponse).
+The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@brick2013Nonresponse).
 The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generated more accurate.

### 2024-11-04 paper/paper.qmd
edits to paper

@@
 despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model Kamala's probability of winning those electoral votes is very high.
-This is due to the small standard deviations shown in @Tab-2 making the predicted outcome of a victory more secure.
+This is due to the small standard deviations shown in @Fig-5 making the predicted outcome of a victory more secure.
 
@@
 
-## 1. Sampling Approach
+## Sampling Approach
 
@@
 
-## 2. Recruitment of Respondents
+## Recruitment of Respondents
 
@@
 
-## 3. Data Validation and Bias Reduction
+## Data Validation and Bias Reduction
 
@@
 
-## 4. Poll Aggregation
+## Poll Aggregation
 
@@
 
-## 5. Budget Allocation
+## Budget Allocation
 
@@
 
-## 6. Question Design and Goal Alignment
+## Question Design and Goal Alignment
 The goal of our survey is to predict who will win the election, this is the research question that guides our survey.

### 2024-11-04 paper/paper.qmd
minor edits

--- 2024-11-04 paper/paper.qmd
@@
 ```
-
+\newpage
 # 1 Introduction
@@
 Key threshold of 270 electoral vote mark needed to secure a victory is diplayed.
-This visualization shows the democrats lead with 227 votes and the Republicans with 219 vot secured"
+This visualization shows the democrats lead with 227 votes and the Republicans with 219 votes secured."
 # Define states and electoral votes, ensuring votes are listed in reverse for clarity in plot
@@
 
-plt.figure(figsize=(20, 14))
-plt.step(x, cumulative_votes, where='post', color='navy', linewidth=2, marker='o')
+plt.figure(figsize=(18, 14))
+stop = 320
+plt.step(x[:stop], cumulative_votes[:stop], where='post', color='navy', linewidth=2, marker='o')
 x_reversed = np.flip(x)
@@
 # Adding threshold line and markers
-plt.xticks(range(len(states)), states, rotation=90, fontsize=8)
+plt.xticks(range(len(states)), states, rotation=90, fontsize=13)
 plt.axhline(270, color='purple', linestyle='--', linewidth=2)
 plt.text(len(states) / 2, 275, '270 votes to win', horizontalalignment='center', color='purple', fontsize=12)
-plt.xlabel('Electoral College by State', fontsize=12)
-plt.ylabel('Total Votes', fontsize=12)
+plt.xlabel('Electoral College by State', fontsize=15)
+plt.ylabel('Total Votes', fontsize=15)
 plt.title('Cumulative Electoral Votes by State', fontsize=25)
@@
 This is clearly a factor here as the response rate was below 2%.
-The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@brick2013Nonresponse).
+The inclusion of response rate adjustments accounts for the likelihood that certain groups are more or less likely to respond to surveys, and is an efficient way to counter these issues (@brick2013nonresponse).
 The stratification also allows for a more representative expectation from the poll ensuring all significant groups are accurately represented in the sample as they are in the population (@Kish1965Survey), Making the means generated more accurate.
@@
 
-## what is included in the poll?
-
-## Methodology and why?
-
+\newpage
 # References

### 2024-11-04 paper/paper.qmd

@@
 despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model Kamala's probability of winning those electoral votes is very high.
-This is due to the small standard deviations shown in @Tab-2 making the predicted outcome of a victory more secure.
+This is due to the small standard deviations shown in @Fig-4 making the predicted outcome of a victory more secure.
 

### 2024-11-04 paper/paper.qmd
Add survey to paper
final ediing of papr and file structure, update readme and requirements

@@
 This analysis provides insight into electoral dynamics, highlighting how data-driven approaches can clarify potential election outcomes and inform strategic political decision-making.
+thanks: "Code and data supporting this analysis is available at: https://github.com/JfpGilbert0/nov_2024_trump_v_harris"
 ---
@@
 Each entry in the dataset corresponds to a specific question about voting preference in a poll conducted by different polling organizations, which measures the percentage of respondents expressing their support for each candidate in their respective state.
-the surveys utilized one or multiple of the following methods for response gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
-The diversity of the methods of collection supplies a colourful array of polls capturing all facets of voters.
+The surveys utilized one or multiple of the following methods for response gathering: Online Ad, Online Panel, Live Phone, text-to-web, mail-to-web, email, Probability panel, IVR.
+The diversity of the methods of collection supplies a colorful array of polls capturing all facets of voters.
 
@@
 # 2.2 The Dataset
-A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President".
+A swing state in the context of a US election is defined as "a state where the number of Democratic and Republican voters is about the same, that has an important influence on the result of the election of the United States President" (@oxfordlanguages).
 States that do not fall under this definition are not likely to "swing" to another party as we approach the election, and thus their outcome can be reasonably assumed.
@@
 Thus our analysis restricts the datasets to state-level polls in the following states: Arizona, Pennsylvania, North Carolina, Georgia, Nevada, Michigan, and Wisconsin; determined to be swing states in 2024 [@nyt_swing, @FiveThirtyEight].
-In @Fig-1 we see the current distribution of electoral votes that each party hold according to those sources as of October 20th, with the remaining swing states, seen in the middle of this graph set to determine if either Harris or Trump receives over 270th electoral votes.
+In @Fig-1 we see the current distribution of electoral votes that each party holds according to those sources as of October 20th, with the remaining swing states, seen in the middle of this graph, set to determine if either Harris or Trump receives over 270th electoral votes.
 
@@
 #| label: Fig-1
-#| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, accordng to 538.
+#| fig-cap: "This figure shows the cumulative total of electoral votes by state, grouped by current party affiliation, according to 538.
 The chart distinguishes between Democrat, Republican, and swing states, with a color-coded fill under each party’s cumulative curve.
-Key threshold of 270 electoral vote mark needed to secure a victory is diplayed.
+Key threshold of 270 electoral vote marks needed to secure a victory is displayed.
 This visualization shows the democrats lead with 227 votes and the Republicans with 219 votes secured."
@@
 
+stop_index_forward = np.argmax(cumulative_votes >= 320)
+stop_index_reverse = np.argmax(cumulative_votes_reversed >= 312)
+
 plt.figure(figsize=(18, 14))
-stop = 320
-plt.step(x[:stop], cumulative_votes[:stop], where='post', color='navy', linewidth=2, marker='o')
+plt.step(x[:stop_index_forward+1], cumulative_votes[:stop_index_forward+1], where='post', color='navy', linewidth=2, marker='o')
 x_reversed = np.flip(x)
-plt.step(x_reversed, cumulative_votes_reversed, where='post', color='red', linewidth=2, marker='o')
+plt.step(x_reversed[:stop_index_reverse + 1], cumulative_votes_reversed[:stop_index_reverse + 1], where='post', color='red', linewidth=2, marker='o')
 
@@
 # Adding threshold line and markers
-plt.xticks(range(len(states)), states, rotation=90, fontsize=13)
+plt.xticks(range(len(states)), states, rotation=90, fontsize=15)
 plt.axhline(270, color='purple', linestyle='--', linewidth=2)
@@
 By selecting polls from after September, we capture data from a period when voter opinions are more stable and better represent the current state of the race.
+All data was gathered on October 24th 2024 and thus all data and analysis is based on what the polling data reflects at this point in time.
+@Fig-2 show a general overview of the polling data.
 
+\newpage
 ```{python}
-
 #| echo: false
-#| label: Fig-2
-#| fig-cap: "Summary statistics of key sample variabls by state. Shows high sampl size and numeric grade among all polls of swing states."
 
@@
 print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='simple_grid'))
+
+```
+```{python}
+#| echo: false
+#| label: Fig-2
+# This code block exits for the use of self referencing
+print(tabulate(rows, headers=['State', 'Variable', 'Mean', 'Std', 'Min', 'Max', 'Count'], tablefmt='simple_grid'))
 ```
@@
 Presidential Election outcome, the primary variables of interest are derived from aggregated state-level polling data.
-The percentage of respondents favouring a candidate is the key variable used for prediction.
+The percentage of respondents favoring a candidate is the key variable used for prediction.
 Within the collection of polls in our sample, the independent variable `trump_pct` represents the percentage of respondents who indicate support for Donald Trump, and `harris_pct` respectively.
@@
 Polls are taken towards the build-up of an election and thus are taken by respondents at different times.
-@Fig-2 shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
+@Fig-3 shows how the sample of polls are distributed over time in which it is clear that we have a collection of polls from a wide range of periods.
 Because sentiment changes over time we adjust the weight that we place on polls that are closer to the election as these are more indicative of how respondents will be voting on the day.
@@
 The larger the sample size the more the response reflects the true population of the state, this is known as the law of large numbers.
-In @Fig-3 we can see the variety within the state polls.
+In @Fig-2 we can see the variety within the state polls.
 For use in the model, the percentages in these polls are weighted for recency and size of the sampled population.
@@
 
-1.	Polling Bias: The model is contingent upon the quality of the polling data.
+1.
+Polling Bias: The model is contingent upon the quality of the polling data.
 Polls with lower numeric grades or transparency scores are excluded, potentially introducing bias if high-quality polls are not representative of the entire electorate.
@@
 
+\newpage
 # 4 Results {#sec-results}
@@
 #| echo: false
-#| label: Fig-4
 # Data setup
@@
 # Display as a professional table
+print(tabulate(
+    df, headers='keys', tablefmt='grid', showindex=False,
+    floatfmt=(".2f", ".2f", ".2f", ".2f", ".2f", ".2f")
+))
+```
+```{python}
+#| echo: false
+#| label: Fig-4
+# This code block exits for the use of self referencing
 print(tabulate(
@@
 The bars extend to the right or left from the center line at 50%, with the distance representing the magnitude of deviation from an even chance.
-This chart visualizes Trump’s relative strength in each swing state and highlights that he s only favoured in Georgia and Arizona."
+This chart visualizes Trump’s relative strength in each swing state and highlights that he is only favored in Georgia and Arizona."
 # Data setup (can adjust based on model results)
@@
 Kamala is predicted to win the remaining states and is given fairly high chances of locking these up.
-despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model Kamala's probability of winning those electoral votes is very high.
+Despite only winning Nevada, North Carolina and Pennsylvania by under a percentage point in the model, Kamala's probability of winning those electoral votes is very high.
 This is due to the small standard deviations shown in @Fig-4 making the predicted outcome of a victory more secure.
@@
 Taking The probabilities from the model above we can look at the outcomes of the 207 possible combinations of wins and losses for Trump in these swing states and deduce his overall probability of reaching 270 electoral votes.
-Based on the modelled probabilities his current expected vote is 41.05, below the threshold of 51 he needs based on our assumptions.
+Based on the modeled probabilities his current expected vote is 41.05, below the threshold of 51 he needs based on our assumptions.
 
@@
 
-The result of the model tells us that overall the polls in October are favouring Harris to win the election.
+The result of the model tells us that overall the polls in October are favoring Harris to win the election.
 From the perspective of the republican party if they wish to "turn the tide" this model highlights which of the swing states are more likely to flip than others.
@@
 It would be recommended to allocate resources here to try and win a state from Harris.
-As far as retaining currently favoured states, Michigan is in a weak position for Trump according to the model, nearing a coin flip.
+As far as retaining currently favored states, Michigan is in a weak position for Trump according to the model, nearing a coin flip.
 As this state holds 15 electoral College votes, the 4th most of the swing states, it is a key state to win.
-Assuming the model's correct prediction that Michigan is currently held by Trump, expect the Democrats to attempt to tip the scales in their favour there.
+Assuming the model's correct prediction that Michigan is currently held by Trump, expect the Democrats to attempt to tip the scales in their favor there.
+Although FiveThirtyEight defines them as swing states, this model predicts Wisconsin and Arizona are fairly secure for Kamala and Trump respectively.
+Showing how modeling polling data is useful in predicting state results that were previously deemed uncertain.
 
-One of the most significant insights from this analysis is the discrepancy between polling data and the betting markets.
-While the majority of reputable polls suggest that Kamala Harris is favoured on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
+One interesting insight from this analysis is the discrepancy between polling data and the betting markets.
+While the majority of reputable polls suggest that Kamala Harris is favored on both a national level and in most swing states, betting markets like Polymarket imply a 67% chance of Trump winning the election as of October 23, 2024 [@polymarket].
 This stark disconnect raises questions about the accuracy of traditional polling methods and whether betting markets, which are financial tools driven by market forces, may offer a more precise reflection of public sentiment.
@@
 Another important consideration is the possible biases in both polling and betting markets.
-Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favour of Harris.
-Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favour him.
+Poll respondents may not be a representative sample of the electorate; for example, Democrats could be more likely to respond to polls, skewing the results in favor of Harris.
+Meanwhile, individuals who participate in betting markets might form a subset of the population that is disproportionately supportive of Trump, which could explain why the implied odds heavily favor him.
 Furthermore, the rapid expansion and increased liquidity of betting markets in the last few years may have improved their efficiency, making them more reflective of real-world probabilities.
@@
 First, the model is based on polling data available far in advance of the election, meaning there is a high degree of uncertainty, and the model may not fully capture late shifts in public opinion or external shocks (e.g., economic downturns or scandals).
-As polling results are based on public opinion, large shock events are expected, furthermore, scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modelled here.
-Additionally, the use of linear modelling may oversimplify the complex dynamics of voter behaviour, as elections often involve nonlinear influences that are difficult to predict.
+As polling results are based on public opinion, large shock events are expected, furthermore, scheduled events such as debates and rallies can be expected to have large impacts on the election that cannot be modeled here.
+Additionally, the use of linear modeling may oversimplify the complex dynamics of voter behavior, as elections often involve nonlinear influences that are difficult to predict.
 
@@
 Presidential Election, there is still much to learn.
-By refining the model and incorporating additional data, we can move closer to producing predictions that more accurately reflect voter behaviour and election outcomes.
+By refining the model and incorporating additional data, we can move closer to producing predictions that more accurately reflect voter behavior and election outcomes.
 
-# Appendix A: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting
+\newpage
+# Appendix A: Idealized Methodology and Survey for Forecasting the US Presidential Election with Incentivized Betting {#sec-apx-A}
 
@@
 The purpose of this is twofold, it aims to minimize the attrition of the respondents, so that more people complete the survey, and it also will provide us insight into who people think will win rather than just who people want to win.
-Ideally, the purpose of this is so that people take into consideration who their friends, family, and their community, are expecting to win.
+Ideally, the purpose of this is so that people take into consideration who their friends, family, and their community are expecting to win.
 
@@
 
-# Appendix B: Siena / New York Tims Pennsylvania poll
+\newpage
+
+# Appendix B: Siena / New York Tims Pennsylvania poll {#sec-apx-B}
 
@@
 Records are selected by state and sampling is separated for Pennsylvania as a whole and its major city Philadelphia.
-The voter file is then ratified and weights are based on historically modelled response rate and coverage.
+The voter file is then ratified and weights are based on historically modeled response rate and coverage.
 

### 2024-11-04 scripts/xandermodel1.py

@@ -2,7 +2,7 @@ import pandas as pd
 from sklearn.linear_model import LogisticRegression
 
 # Load the cleaned polling data
-clean_poll_data = pd.read_csv("data/02-analysis_data/merged_swing_state_data.csv")
+clean_poll_data = pd.read_parquet("data/02-analysis_data/merged_swing_state_data.parquet")
 
 # Create a weight for each poll based on recency and sample size
 current_date = pd.to_datetime('today')
